### 2022-03-23 analysis/paper/shceof.Rmd
Revisión

---
title: "Revisiting the Austral Spring Extratropical Southern Hemisphere zonally asymmetric circulation using complex Empirical Orthogonal Functions"
titlerunning:  Title
authors: 
- name: Elio Campitelli
  address: |
    Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Ciencias de la Atmósfera y los Océanos. Buenos Aires, Argentina.  
    CONICET – Universidad de Buenos Aires. Centro de Investigaciones del Mar y la Atmósfera (CIMA). Buenos Aires, Argentina.  
    CNRS – IRD – CONICET – UBA. Instituto Franco-Argentino para el Estudio del Clima y sus Impactos (IRL 3351 IFAECI). Buenos Aires, Argentina.
  email: elio.campitelli@cima.fcen.uba.ar

- name:  Leandro B. Díaz
  address: |
    Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Ciencias de la Atmósfera y los Océanos. Buenos Aires, Argentina.
    CONICET – Universidad de Buenos Aires. Centro de Investigaciones del Mar y la Atmósfera (CIMA). Buenos Aires, Argentina.
    CNRS – IRD – CONICET – UBA. Instituto Franco-Argentino para el Estudio del Clima y sus Impactos (IRL 3351 IFAECI). Buenos Aires, Argentina.

- name:  Carolina Vera
  address: |
    Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Ciencias de la Atmósfera y los Océanos. Buenos Aires, Argentina.
    CONICET – Universidad de Buenos Aires. Centro de Investigaciones del Mar y la Atmósfera (CIMA). Buenos Aires, Argentina.
    CNRS – IRD – CONICET – UBA. Instituto Franco-Argentino para el Estudio del Clima y sus Impactos (IRL 3351 IFAECI). Buenos Aires, Argentina.

thanks: | 
  The research was supported by UBACyT20020170100428BA and the CLIMAX Project funded by Belmont Forum/ANR-15-JCL/-0002-01. Elio Campitelli was supported by a PhD grant from CONICET, Argentina.

keywords:

abstract: |
  abstract

date: "`r format(Sys.time(), '%d %B, %Y')`"

bibliography: 
  - references.bib
  - data.bib
  - packages.bib

bibstyle: spbasic
# bibstyle options spbasic(default), spphys, spmpsci
output: 
  # bookdown::word_document2: default
  bookdown::pdf_book: 
      base_format: rticles::springer_article
      latex_engine: "xelatex"
      # citation_package: "natbib"
      # pandoc_args: ["--lua-filter=abstract-to-meta.lua"]
always_allow_html: true
header-includes: 
 - \usepackage{gensymb}
 - \usepackage{subfig}
 - \usepackage[inline]{showlabels}
 - \usepackage{chngcntr}
 - \usepackage{natbib}
 - \usepackage{lineno}
 - \linenumbers 
---

```{r setup, echo = FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  cache = TRUE,
  comment = "#>",
  cache.extra = 41,
  cache.path = "../cache/paper/",
  fig.path = "../figures/",
  dpi = 300,
  dev = c("pdf", "png")
)

width_column <- 3.3

library(metR)
library(data.table)
library(magrittr)
library(ggplot2)
library(tagger)
library(ggperiodic)
library(patchwork)
library(kableExtra)
library(shceof)

theme_set(theme_shceof(base_size = 10) +
            theme(legend.box.spacing = grid::unit(-.5, "lines")))

panel_background <- theme(panel.background = element_rect(fill = "#fbfbfb", color = NA), 
                          panel.ontop = FALSE,
                          tagger.panel.tag.background = ggplot2::element_rect(color = NA, 
                                                                              fill = "#fbfbfb"))

grid <- theme(panel.grid = element_line(color = "gray10", size = 0.4, linetype = 3))
no_grid <- theme(panel.grid = element_blank())
grid_y <- theme(panel.grid.major.y = element_line(color = "gray60", size = 0.1))

guide_colorstrip_bottom <- function(width = 15, height = 0.5, 
                                    title.position = "top",...) {
  guide_colorstrip(title.position = title.position, title.hjust = 0.5,
                   barheight = height,
                   barwidth = width, ...)
}

guide_colorsteps <- function (even.steps = TRUE, show.limits = NULL, ticks = FALSE, 
                              ...) {
  guide <- guide_colourbar(raster = FALSE, ticks = ticks, 
                           nbin = 300, ...)
  guide$even.steps <- even.steps
  guide$show.limits <- show.limits
  class(guide) <- c("colorsteps", class(guide))
  guide
}
guide_colorsteps_bottom <- function(width = 15, height = 0.5, even.steps = FALSE, 
                                    title.position = "top", show.limits = TRUE, ticks = TRUE,
                                    ...) {
  guide_colorsteps(show.limits = show.limits, even.steps = even.steps,
                   title.position = title.position, title.hjust = 0.5,
                   barheight = height, ticks = ticks,
                   ticks.colour = "black",
                   barwidth = width, frame.colour = "black", ...)
}

axis_labs_smol <- theme(axis.text = element_text(size = 6))

pattern_dots <- ggplot2::ggproto("GeomDots", ggpattern::GeomPolygonPattern)
ggplot2::update_geom_defaults(pattern_dots, 
                              list(pattern = "circle",
                                   colour = NA,
                                   pattern_colour = NA,
                                   pattern_fill = "black",
                                   pattern_density = 0.3,
                                   pattern_alpha = 1,
                                   pattern_spacing = 0.02,
                                   fill = NA
                              ))
update_geom_defaults(GeomSmooth, list(colour = "#0d52bf"))

f <- formals(geom_contour_tanaka)
f$range <-  c(0.01, 0.2)
formals(geom_contour_tanaka) <- f

geom_contour_pval <- function(mapping, p.value = 0.01, size = 0.1, ...) {
  mapping2 <- mapping
  mapping2$fill <- aes(fill = NA)$fill
  list(
    stat_contour_filled(mapping2, breaks = c(0, p.value), fill = NA, geom = pattern_dots, ...),
    geom_contour2(mapping, breaks = p.value, size = size, ...)
  )
}

# There's a bunch of issues with s2 right now
sink <- capture.output(sf::sf_use_s2(FALSE))

geom_coords <- function() {

  list(
    no_grid,
    annotate("segment",
             y = seq(-90, 0, by = 15),
             yend = seq(-90, 0, by = 15),
             x = 0,
             xend = 360,
             size = 0.1, alpha = 0.5),

    annotate("segment",
             x = seq(0, 360 - 30, by = 30),
             xend =  seq(0, 360 - 30, by = 30),
             y = -90 + 15,
             yend = Inf,
             size = 0.1, alpha = 0.5),
    shadowtext::geom_shadowtext(data = data.frame(x = 0, y = seq(-90 + 15, 0, 
                                                                 by = 15)),
                                aes(x, y, label = LatLabel(y)), size = 1.5, 
                                alpha = 0.7,
                                colour = "black",
                                bg.colour = "white")
  )
}

coord_polar <- function(ymax = -20, ...) {
  x <- c(seq(0, 360, length.out = 40), 
         seq(360, 0, length.out = 40), 
         0)
  y <- c(rep(ymax, length.out = 40), 
         rep(60, length.out = 40), 
         ymax)

  cbind(x, y) %>% 
    list() %>% 
    sf::st_polygon() %>% 
    sf::st_sfc(crs = "+proj=latlong") -> white

  list(
    geom_sf(data = white, inherit.aes = FALSE, 
            fill = "white", 
            colour = "white", size = 2),
    coord_sf(ylim = c(-90, ymax), 
             lims_method = "box",
             crs = "+proj=laea +lat_0=-90",
             default_crs = "+proj=longlat",
             label_axes =  "----", ...)
  )
}

one_less <- function(x) {
  n <- length(x)
  if (n < 4) return(x)
  x <- round(x, 5)

  c( rev(JumpBy(rev(x[x < 0]), 2, fill = "")),
     JumpBy(x[x > 0], 2, fill = ""))
}

AnchorBreaks <- function (anchor = 0, binwidth = NULL, exclude = NULL, bins = 10, sym = FALSE) {
  force(anchor)
  force(binwidth)
  force(exclude)
  force(bins)
  force(sym)
  function(x, binwidth2 = NULL) {
    if (sym) {
      length <- max(abs(x - anchor))
      x <- c(anchor - length, anchor + length)
    }
    if (!is.null(binwidth)) 
      binwidth2 <- binwidth
    if (is.null(binwidth2)) {
      binwidth2 <- diff(pretty(x, bins))[1]
    }
    mult <- ceiling((x[1] - anchor)/binwidth2) - 1L
    start <- anchor + mult * binwidth2
    b <- seq(start, x[2] + binwidth2, binwidth2)
    b[!(b %in% exclude)]
  }
}

lab_lev <- AddSuffix(" hPa")
lab_cplx <- c(R = "Real", 
              I = "Imaginary")

combine_lists <- function(...) {
  # browser(expr = length(x) != length(y))
  X <- list(...)
  names <- names(X[[1]])
  names(names) <- names
  lapply(names, function(n) {
    unlist(lapply(X, function(x) x[[n]]))
  })
}

correlation_statistics <- function(x, n, signif = 2) {
  z <- atanh(x)
  se <- 1/sqrt(n - 3)

  low <- qnorm(.025, mean = z, sd = se)
  low <- tanh(low)
  hig <- qnorm(.975, mean = z, sd = se)
  hig <- tanh(hig)

  p.value <- pnorm(abs(x)/se, lower.tail = FALSE)

  text <- paste0(signif(x, signif), " (CI: ",
                 signif(low, signif), " -- ", signif(hig, signif), ")")

  list(estimate = x, 
       p.value = p.value, 
       low = low,
       hig = hig,
       text = text
  )
}

```

```{r read-indices}
enso <- rsoi::download_oni(TRUE, data_path("raw", "oni.csv")) %>%
  as.data.table() %>%
  .[, .(time = lubridate::as_datetime(Date), oni = ONI)] %>%
  na.omit() %>%
  .[season(time) == "SON"] %>%
  .[, oni := oni*season_weights(time)] %>% 
  .[, .(oni = mean(oni)), by = .(time = seasonally(time))]

dmi <- rsoi::download_dmi(TRUE, data_path("raw", "dmi.csv")) %>%
  as.data.table() %>%
  .[, .(time = lubridate::as_datetime(Date), dmi = DMI)] %>%
  na.omit() %>%
  .[season(time) == "SON"] %>%
  .[, dmi := dmi*season_weights(time)] %>% 
  .[, .(dmi = mean(dmi)), by = .(time = seasonally(time))]

sams <- readRDS(data_path("derived", "sam.Rds")) %>%
  .[season(time) == "SON"] %>%
  .[, estimate := estimate*season_weights(time)] %>% 
  .[, .(estimate = mean(estimate)),
    by = .(lev, term, time = seasonally(time))]

levs <- unique(sams$lev)

for (l in seq_along(levs)[-1]) {
  cor <- sams[lev %in% levs[seq(l - 1, l)]] %>% 
    .[term == "full"] %>%
    dcast(time ~ lev, value.var = "estimate") %>% 
    setnames(., colnames(.), c("time", "old", "new")) %>% 
    .[, cor(old, new)]

  sams[lev == levs[l], estimate := sign(cor)*estimate]
}
```

```{r test-irlba}
x <- "irlba"
has_irlba <- requireNamespace(x, quietly = TRUE)
if (has_irlba) {
  stop("If installed, metR::EOF uses irlba, but that package has a bug with complex matrices\n(see: https://github.com/bwlewis/irlba/issues/55). Uninstall irlba with renv::remove(\"irlba\") and try again.")
}

```

```{r read-data}
era5 <- ERA5_geopotential() %>% 
  ReadNetCDF(vars = c(hgt = "z"),
             subset = list(level = list(200, 50),
                           latitude = -90:10)) %>% 
  normalise_coords() %>% 
  .[season(time) == "SON"] %>%
  .[, w := season_weights(time)] %>% 
  .[, .(hgt = mean(hgt*w)), by = .(lev, lon, lat, time = seasonally(time))]
```

```{r compute-psi}
# Scaling magical numbers. (come from psi-test.Rmd)
# Not really important, though.
scaling <- c(add = -2.364646e+06, mult = 3.954828e+13)
psi <- ERA5_vorticity() %>% 
 ReadNetCDF(vars = "vo") %>%
  normalise_coords() %>%
  .[, lev := 200] %>% 
  .[season(time) == "SON"] %>%
  .[, w := season_weights(time)] %>% 
  .[, .(vo = mean(vo*w)), by = .(lon, lat, lev, time = seasonally(time))] %>%
  .[, solve_poisson(vo, lon, lat), by = .(lev, time)] %>%
  setnames("value", "psi") %>% 
  .[, psi := psi*scaling[2] + scaling[1]] %>%
  .[, psi.z := Anomaly(psi), by = .(time, lev, lat)] 
```

```{r o3}
o3 <- ERA5_TOC() %>% 
  ReadNetCDF(vars = c(toc = "tco3")) %>% 
  normalise_coords() %>% 
  .[lat <= -20] %>% 
  .[, toc := toc/2.1415e-5] %>%    # transform to Dobson Units (https://sacs.aeronomie.be/info/dobson.php)
  .[, w := season_weights(time)] %>% 
  .[, .(toc = mean(toc*w)), by = .(lon, lat, time = seasonally(time))]

o3wave_lats <- c(-75, -45)
```

```{r psa}
psa <- ERA5_geopotential() %>% 
  ReadNetCDF(vars = c(hgt = "z"),
             subset = list(level = list(500),
                           latitude = -90:0)) %>% 
  normalise_coords() %>% 
  .[, w := season_weights(time)] %>% 
  .[, .(hgt = mean(hgt*w)), by = .(lon, lat, time = seasonally(time))] %>% 
  .[, value := Anomaly(hgt)*sqrt(cos(lat*pi/180)), 
    by = .(lon, lat, season(time))] %>%
  EOF(value ~ time | lon + lat, n = 1:3, data = .)

psa <- lapply(psa, function(x) {
  copy(x)[, PC := fcase(PC == "PC1", "SAM", 
                        PC == "PC2", "PSA1", 
                        PC == "PC3", "PSA2")] 
})

psa_time <- psa$left[season(time) == "SON"]
```

# Introduction

The large-scale extratropical circulation in the Southern Hemisphere (SH) is strongly zonally symmetric, but its zonal departures are highly relevant for regional impacts [e.g. @hoskins2005].
They strongly modulate weather systems and regional climate through promoting longitudinally different latitudinal transport of heat, humidity, and momentum [@trenberth1980a; @raphael2007] and could even be related to the occurrence of high-impact climate extremes [@pezza2012].

Zonally asymmetric circulation is typically described by the amplitude and phase of zonal waves obtained by Fourier decomposition of geopotential heights or sea-level pressures at each latitude [e.g. @vanloon1972; @trenberth1980a; @turner2017].
This approach suggests that zonal waves 1 and 3 explain almost 99% of the total variance in the annual mean 500-hPa pattern at 50ºS [@vanloon1972].
@trenberth1985 concluded that wave 3 plays a role in the development of blocking events.
In addition, previous works have identified at extratropical and subpolar latitudes, wave-like patterns with dominant wavenumbers 3-4, also exerting distinctive regional impacts.
@raphael2007 shows that variability in the planetary wave 3 projected onto its climatological location is associated with anomalies in the Antarctic sea-ice concentration.

The Fourier decomposition relies on the assumption that the circulation can be meaningfully described in terms of zonal waves of constant amplitude along a latitude circle.
However, this is not valid for meridionally propagating waves or zonal waves with localised amplitudes.
Addressing this limitation, the Fourier technique can be generalized to integrate all planetary wave amplitude regardless of wave number, by computing the wave envelope [@irving2015].
The latter makes it possible to represent planetary waves with different amplitude at different longitudes, but it removes all information about phase and wave number.
With this approach, @irving2015 showed that planetary wave amplitude in general is associated to Antarctic sea-ice concentration and temperature, as well as to precipitation anomalies in regions of significant topography in SH mid-latitudes and Antarctica.

Another extensively used approach to characterize the SH tropospheric circulation anomalies, is by computing Empirical Orthogonal Functions (EOF, also known as Principal Component Analysis).
Within the EOF framework, the Southern Annular Mode (SAM) appears as the leading mode of variability of the SH circulation [@fogt2020].
SAM represents a relatively zonally symmetric pattern of alternating low pressures in polar latitude and a ring of high pressures in high latitudes with an embedded wave 3 pattern that is more prominent in the Pacific sector.
The 2nd and 3rd EOFs, usually known as Pacific--South American Patterns (PSA) 1 and PSA2 patterns, respectively, describe meridionally propagating wave trains that originate in the eastern equatorial Pacific and Australian-Indian Ocean sector, and travel towards the South Atlantic following a great-circle arch along the Antarctic coast [@mo2001].
These patterns influence precipitation anomalies in South America [@mo2001].
Although these patterns are usually derived by applying EOF to temporal anomalies, [@raphael2003] also applied EOF methods specifically to zonal anomalies.
More recently, [@irving2016] proposed a novel methodology for objectively identifying the PSA pattern using Fourier decomposition.

Patterns resulting from EOF analyses are more flexible than Fourier decomposition derived modes in the sense that they can capture oscillation patterns that cannot be characterised by purely sinusoidal waves with constant amplitude.
Nonetheless, they are restricted to standing oscillation modes and could not represent properly propagating or phase-varying features such as zonal waves.
A single EOF can also represent a mixture of two or more physical modes.

A third methodology commonly used consists on identifying particular features of interest and creating indices using simple methods such as averages and differences.
Examples of this methodology are the SAM Index of [@gong1999], the SH wave 3 activity index defined by @raphael2004 and the SH zonally asymmetric circulation index from @hobbs2010.
These derived methods are grounded on other methods such as Fourier decomposition or EOF to identify the centres of action for the described phenomena and can be useful to characterise features that are not readily apparent with these methods.
It is commonly easy the computation of these kind of indices, but they could be unable to capture non-stationary patterns.

An alternative methodology that has been proposed to study travelling and standing waves is complex Empirical Orthogonal Functions [@horel1984].
This method extends EOF analysis to capture oscillations with varying amplitude and phase and has been applied to the time domain.
For instance, @krokhin2007 applied cEOF to station-based monthly precipitation anomalies and monthly temperature anomalies in the Eastern Siberia and the Far East region to characterise the main modes of variability and their connection to teleconnection indices.
Similarly, @gelbrecht2018 used cEOF applied to daily precipitation from reanalysis to study the propagating characteristics of the South American Monsoon.
To our knowledge, complex EOF has not been applied in the spatial domain to capture the phase-varying nature of planetary waves.

Some of the zonally asymmetric patterns of the SH circulation variability described previously, appear to have experienced secular changes.
For instance, @raphael2003 found that the amplitude of the zonal wave 1 experienced a large increase and that the zonal wave 3 experienced changes in its annual cycle between 1958 and 1996.
However, little is known yet about variability and trends of these patterns.

The general goal of this study is to improve the description and understanding of the zonally asymmetric extratropical SH circulation using complex Empirical Orthogonal Functions, which allow to describe phase varying planetary waves with variable amplitudes along a latitude circle.
In addition, it is proposed to expand the knowledge of the simultaneous behaviour of SH asymmetric circulation in the troposphere and the stratosphere.

We restrict this paper to the September-October-November (SON) trimester.
We chose this season because it showed the strongest and more clearly interpretable results in a preliminary analysis of the four season.
This is is consistent with previous research that shows that this season experiences a maximum in tropical teleconnections [@cazes-boezio2003a] and stratosphere-troposphere interactions [@lim2018].

In Section \@ref(methods) we describe the methods.
In Section \@ref(spatial) we analyse the spatial patterns of each complex EOF.
In Section \@ref(regressions) we study the spatial regressions with geopotential height and temperature anomalies.
In Section \@ref(other-variables) we analyse the relationship between cEOF1 and Ozone, and between cEOF2 with the SAM and the PSA.
In Section \@ref(tropical) we study tropical forcings that explain the variability of each cEOF.
Finally, in Section \@ref(precipitation) we show the relationship between these modes of variability and precipitation anomalies in South America and Oceania.

<!-- The large-scale extratropical circulation in the Southern Hemisphere (SH) is strongly zonally symmetric, but its zonal departures are highly relevant for regional impacts and could even be related to the occurrence of high-impact climate extremes (REFERENCE).  -->

<!-- Zonal asymmetries of extratropical circulation in the Southern Hemisphere strongly modulate weather systems and regional climate through promoting longitudinally different latitudinal transport of heat, humidity, and momentum [@trenberth1980]. -->

<!-- Hablar simplemente de cómo estos fenómenos varían en longitud.  -->

<!-- @hill2009 ->  -->

<!-- Comparan eventos ENSO del 97 y de 2002 -->

<!-- La posición de las anomalías de circulación (PSA) son importantes para determinar el impacto del ENSO en sudamérica. Sugiere que la localización de la anomalía de SST cambia la localización del PSA.  -->

<!-- > The meridional winds from the idealised experiments based on simplified and localised SST warming showed a significant difference between experiments forced by eastern versus central-west Pacific SST anomalies, with a clear phase shift of the PSA apparent. This phase shift in the idealised experiments is possibly the main factor responsible for the different responses in South American atmospheric circulation between the 1997/1998 and 2002/2003 El Niño events. The phase shift over southeastern South America was also seen in the low-level 850 hPa winds in the EN97 and EN02 experiments, suggesting that the PSA phase shift is responsible for the shift in the moisture transport over southeastern Brazil between the two El Niño events. In summary, the location of the maximum SST anomaly appears to be an important factor in controlling the amount of moisture transported into the SACZ region during El Niño events.  -->

<!-- @hobbs2010 -> muestra la importancia de la longitud de ciclones -->

<!-- ciasto2015 -> no investiga impactos -->

<!-- sun2013 -->

<!-- wilson2014 -> -->

<!-- Observaciones y modelos enso canonico y modoki. diferentes patrones de circulacion.  -->

<!-- wang2021 ->  -->

<!-- > However, the spatial pattern of the PSA wave train excited by ENSO seems to be -->

<!-- > non-stationary. Fogt and Bromwich (2006) indicated that the PSA mode moved -->

<!-- > eastwards by 20° longitude in the 1990s owing to the eastward shift of the tropical -->

<!-- > convective position. In addition, Wilson et al. (2014, 2016) and Ciasto et al. (2015) -->

<!-- > found that the different types of ENSO can contribute to the wave shift. Specifically, -->

<!-- > Rossby waves excited by central Pacific (CP)–type El Niño shift westwards owing to -->

<!-- > the westward movement of the tropical diabatic heating. -->

<!-- Diferencia en la localización de los centros en LN y EN -> diferencias en los impactos en la península antártica.  -->

<!-- Lo más importante es la localización de la salida del jet, no la perturbación de sst.  -->

<!-- fogt2006 ->  -->

<!-- cambios en la posición de las teleccionexiones. Pero es todo con composiciones y comparando décadas. No mira impactos. -->

<!-- lou2021 ->  -->

<!-- > In combination, they produce the single propagating PSA mode. -->

<!-- Complicado y no entiendo mucho, pero cosas que puedo ver es que PSA forza el océano subtropical que luego es precursora de ENSO. No sé si eso es súper  -->

# Data and Methods {#methods}

## Data

We used monthly geopotential height, air temperature, ozone mixing ratio, and total ozone column (TOC) at 2.5\degree longitude by 2.5\degree latitude of horizontal resolution and 37 vertical isobaric levels from the European Centre for Medium-Range Weather Forecasts Reanalysis version 5 [ERA; @era5] for the period 1979 -- 2019.
Most of our analysis is restricted to the post-satellite era to avoid any confounding factors arising from the incorporation of satellite observations, but we also used the preliminary back extension of ERA5 from 1950 to 1978 [@era5be] to describe long-term trends.
We derived streamfunction at 200 hPa from ERA5 vorticity using the FORTRAN subroutine FISHPACK [@fishpack] and we computed horizontal wave activity fluxes following @plumb1985.
Sea Surface Temperature (SST) monthly fields are from Extended Reconstructed Sea Surface Temperature (ERSST) v5 [@huang2017] and precipitation monthly data from the CPC Merged Analysis of Precipitation [@cmap], with a 2.5\degree resolution in latitude and longitude.
The rainfall gridded dataset is based on information from different sources such as rain gauge observations, satellite inferred estimations and the NCEP-NCAR reanalysis, and it is available since 1979 to present.

The Oceanic Niño Index [ONI, @bamston1997] comes from NOAA's Climate Prediction Center and the Dipole Mode Index [DMI, @saji2003] from Global Climate Observing System Working Group on Surface Pressure.

## Methods

The study is restricted to the spring season, defined as the September-October-November (SON) trimester.
We compute seasonal means for the different variables, averaging monthly values weighted by the number of days in each month.
We use the 200 hPa level to represent the high troposphere and 50 hPa to represent the lower stratosphere.

The amplitude of the zonal waves was obtained through the Fourier transform of the spatial field at each latitude circle.
For the analysis of wave 1, we computed its amplitude and phase by averaging (area-weighted) the data for each variable and each SON between `r knitr::combine_words(LatLabel(o3wave_lats))`, and then extracting the wave-1 component of the Fourier spectrum.
We chose this latitude band because it is wide enough to capture most of the relevant anomalies of SH mid-latitudes.

We computed the level-dependent SAM index as the leading EOF of year-round monthly geopotential height anomalies south of 20ºS at each level for the whole period [@baldwin2009].
We further split the SAM into its zonally symmetric and zonally asymmetric components (S-SAM and A-SAM indices respectively).
These indices were obtained by projecting the zonally asymmetric and zonally symmetric part of the SAM spatial pattern onto monthly geopotential height fields, as proposed by @campitelli2022.
Seasonal indices of the Pacific South American patterns (PSA1 and PSA2) were calculated, in agreement with @mo2001, as the third and fourth leading EOF of seasonal mean anomalies for 500-hPa geopotential heights at SH.

Linear trends were computed by Ordinary Least Squares (OLS) and the 95% confidence interval was computed assuming a t-distribution with the appropriate residual degrees of freedom [@wilks2011].

## Complex Empirical Orthogonal Functions (cEOF)

(ref:eof-naive-cap) Spatial patterns of the four leading EOFs of SON zonal anomalies of geopotential height at 50 hPa south of 20\degree S for the 1979 -- 2019 period (arbitrary units).

```{r compute-eof-naive}
eof_naive <- era5[lev %in% c(50, 200) & lat %between% c(-90, -20)] %>% 
  # .[, hgt := Anomaly(hgt), by = .(season(time), lon, lat, lev)] %>%
  .[, hgt  := Anomaly(hgt)*sqrt(cos(lat*pi/180)), by = .(time, lev, lat)] %>% 
  .[, .(eof = list(EOF(hgt ~ time | lon + lat, n = 1:4, data = .SD, suffix = "EOF"))), 
    by = lev]

explained <- eof_naive[, eof[[1]]$sdev, by = lev] %>% 
  .[order(lev, EOF)] %>% 
  .[, scales::percent(r2, accuracy = .1)]

```

```{r eof-naive, fig.cap = "(ref:eof-naive-cap)", fig.width=width_column, fig.height=width_column*1.2}
eof_naive[, eof[[1]]$right, by = lev] %>%
  .[lev == 50] %>% 
  .[ , hgt := hgt/sd(hgt)] %>% 
  periodic(lon = c(0, 360)) %>%
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = hgt, fill = ..level..),
                    breaks = AnchorBreaks(0, 1, exclude = 0)) +
  geom_contour_tanaka(aes(z = hgt), breaks = AnchorBreaks(0, 1, exclude = 0), 
                      range = c(0.01, 0.2)) +

  geom_qmap(size = 0.1) +
  scale_fill_divergent_discretised(NULL, guide = guide_colorsteps_bottom(10))  +
  scale_y_latitude(labels = NULL,
                   minor_breaks = NULL, 
                   limits = c(NA, -20)) +
  scale_x_longitude() +
  geom_coords() +
  coord_polar() +
  theme(axis.text = element_blank(), 
        legend.box.spacing = grid::unit(-1, "lines")) +
  facet_wrap(.~EOF, ncol = 2)  +
  tagger::tag_facets("rc")

```

In the standard EOF analysis, zonal waves may appear as pairs of (possibly degenerated) EOFs representing similar patterns but shifted in phase [@horel1984].
Figure \@ref(fig:eof-naive) shows the four leading EOFs of SON geopotential height zonal anomalies at 50 hPa south of 20\degree S.
It is clear that the first two EOFs represent a single phase-varying zonal wave 1 pattern and the last two represent a similarly phase-varying pattern with shorter wavenumber and four centres of action shifted by 1/4 wavelength.
A similar EOF structure can be seen in 200 hPa (not shown).
Since each pair of EOFs seems to represent the same phase-varying structure, it would be desirable to combine them into a single pattern described by amplitude and phase.

Complex Empirical Orthogonal Functions (cEOF) are a useful method for characterising zonal waves, associated with phase-varying structures [@horel1984].
This method involves computing EOF on the analytic representation of the original field.
That representation is a complex field in which the real part is the original data and the imaginary part is the original data shifted by 90\degree at each spectral frequency -- i.e. its Hilbert transform.
The Hilbert transform is usually understood in terms of time-varying signal.
However, in this work we apply the Hilbert transform at each latitude circle and at each considered time.
Since each latitude circle is a periodic domain, this procedure does not suffer from edge effects.

The result of the cEOF methodology is a set of complex spatial patterns and complex time series.
The real and imaginary part of each spatial pattern represent the two phases of wave-like spatial patterns that are in quadrature.
The magnitude and argument of each complex time series represent the amplitude and phase of each zonal wave.

```{r compute-ceof}
invisible(
  ERA5_geopotential() %>% 
    ReadNetCDF(vars = c(hgt = "z"),
               subset = list(latitude = -90:10,
                             level = list(200, 50))) %>%
    normalise_coords() %>%
    .[season(time) == "SON"] %>%
    .[, w := season_weights(time)] %>% 
    .[, .(hgt = mean(hgt*w)), by = .(lev, lon, lat, time = seasonally(time))] %>%
    .[, season := season(time)] %>%
    {
      ceof <<- .[, .(eof = list(compute_ceof(.SD))), by = .(season)]
      ceof_splitted <<-  .[, levs := lev] %>%
        .[, .(eof = list(compute_ceof(.SD, n = 1:3))), by = .(season, levs)]

    }
)
```

(ref:corr-ceof-splitted-cap) Coefficient of determination ($R^2$) between the absolute magnitude of complex EOFs computed separately at 200 hPa and 50 hPa (p-values lower than 0.01 in bold).

```{r corr-ceof-splitted}
table <- ceof_splitted[, eof[[1]]$left, by = .(lev = levs)] %>%
  copy() %>%
  .[, mag  := abs(hgt)] %>%
  .[, item := paste(lev, cEOF, sep = "_")]

n <- uniqueN(table$time)

table2 <- table %>%
  widyr::pairwise_cor(item, time, mag) %>% 
  setDT() %>%
  .[, p.value := correlation_statistics(correlation, n)$p.value] %>% 
  .[tstrsplit(item1, "_")[[1]] != tstrsplit(item2, "_")[[1]], ] %>%
  .[tstrsplit(item1, "_")[[1]] != 50] %>%
  .[, item1 := tstrsplit(item1, "_")[[2]]] %>%
  .[, item2 := tstrsplit(item2, "_")[[2]]] %>%
  .[, r2 := round(correlation^2, 2)]

pval <- table2 %>% 
  dcast(item1 ~ item2, value.var = c("p.value"))

table2 %>% 
  dcast(item1 ~ item2, value.var = c("r2")) %>% 
  setnames("item1", "200 hPa") %>%
  kbl2(booktabs = TRUE, caption = "(ref:corr-ceof-splitted-cap)") %>%
  column_spec2(2:4, bold = pval[[.col]] < 0.01) %>%
  add_header_above(c("", "50 hPa" = 3)) %>%
  kable_classic() 
```

The cEOF methodology is applied to SON geopotential height zonal anomalies south of 20\degree S separately at 50 and 200 hPa.
Table \@ref(tab:corr-ceof-splitted) shows the coefficient of determination between time series of the amplitude of each cEOF across levels.
There is a high degree of correlation between the magnitude of the respective cEOF1 and cEOF2 at each level.
The spatial patterns of the 50 hPa and 200 hPa cEOFs are also similar (not shown).

Both the spatial pattern similarity and the high temporal correlation of cOEFs computed at 50 hPa and 200 hPa suggest that these are, to a large extent, modes of joint variability.
This motivates the decision of performing complex EOF jointly between levels.
Therefore cEOFs were computed using data from both levels at the same time.
In that sense each cEOF has a spatial component that depends on longitude, latitude and level, and a temporal component that depends only on time.

The phase of cEOFs is defined up to an additive constant.
For real EOFs, this constant can be either 0 or $\pi$, corresponding to a change in sign.
For cEOFs, it can be any real number between 0 and $2\pi$ [@horel1984], corresponding to rotations in the complex plane.
Since any choice is arbitrary and equally valid, we chose the phase of each cEOF so that the real or imaginary parts are aligned with meaningful phases in our analysis.

Preliminary analysis showed that the first cEOF was closely related to the the zonal wave 1 of Total Ozone Column and the second cEOF was closely related to ENSO.
Therefore, to aid in the interpretation, we chose the phases of each cEOF so that these relationships are aligned with one of the complex components.

We chose the phase of cEOF1 so that the time series corresponding to the real part has the maximum correlation with the zonal wave 1 of Total Ozone Column between `r knitr::combine_words(LatLabel(o3wave_lats))`.
And we chose the phase of cEOF2 so that the coefficient of determination between the Oceanic Niño Index [@bamston1997] and the real part is minimized, which also nearly maximizes the correlation with the imaginary part.

This procedure does not create spurious correlations, it only takes an existing relationship and aligns it with a specific phase.

```{r rotate_eofs}
rotate <- function(z, angle = 0) {
  complex(real = cos(angle), imaginary = sin(angle)) * z
}

angles <- seq(-pi, pi, by = .5*pi/180)

# rotations_cEOF1 <- lapply(angles, function(a) {
#   ceof$eof[[1]]$left[cEOF == "cEOF1"] %>%
#     copy() %>%
#     .[, hgt := rotate(hgt, a)] %>%
#     sep_ReIm() %>%
#     .[, .(mean = mean(hgt)), by = part] %>%
#     .[, rotation := a]
# }) %>%
#   rbindlist()
# 
# 
# best_rotation_cEOF1 <- rotations_cEOF1[part == "Real"][which.min(mean)]$rotation

wave1_o3 <- o3 %>% 
  .[lat %between% o3wave_lats] %>% 
  .[, .(toc = weighted.mean(toc, w = cos(lat*pi/180), na.rm = TRUE)), by = .(time, lon)] %>% 
  .[, FitWave(toc, 1), by = .(time)]

with_o3 <- ceof$eof[[1]]$left %>% 
  .[cEOF == "cEOF1"] %>% 
  .[wave1_o3, on = "time"]

rotations_o3 <- lapply(angles, function(a) {
  with_o3 %>% 
    copy() %>% 
    .[, hgt := rotate(hgt, a)] %>% 
    na.omit() %>% 
    sep_ReIm() %>% 
    .[, .(cor = cor(amplitude, hgt)), by = part] %>% 
    .[, rotation := a]
}) %>% 
  rbindlist()

best_rotation_cEOF1 <- rotations_o3[part == "Real"][which.max(cor)]$rotation

with_enso <-  cut(ceof$eof[[1]], 2)$left %>%
  copy() %>%
  .[enso, on = "time"] %>% 
  na.omit()

rotations_cEOF2 <- lapply(angles, function(a) {
  with_enso %>%
    .[, hgt2 := rotate(hgt, a)] %>%
    .[, .(R = cor(Re(hgt2), oni),
          I = cor(Im(hgt2), oni))] %>%
    .[, rotation := a]
}) %>%
  rbindlist()

best_rotation_cEOF2 <- rotations_cEOF2[I > 0][which.min(abs(R))]$rotation

eof_rotated_left <- ceof$eof[[1]]$left %>%
  copy() %>%
  .[cEOF == "cEOF2", hgt := rotate(hgt, best_rotation_cEOF2)] %>%
  .[cEOF == "cEOF1", hgt := rotate(hgt, best_rotation_cEOF1)]

eof_rotated_right <-  ceof$eof[[1]]$right %>%
  copy() %>%
  .[cEOF == "cEOF2", hgt := rotate(hgt, best_rotation_cEOF2)] %>%
  .[cEOF == "cEOF1", hgt := rotate(hgt, best_rotation_cEOF1)]

ceof_unrotated <- copy(ceof)

ceof$eof[[1]]$left <- eof_rotated_left
ceof$eof[[1]]$right <- eof_rotated_right

ceof <- rbindlist(list(rotated = ceof,
                       unrotated = ceof_unrotated),
                  idcol = "rotation")

rotated <- function(x) x[rotation == "rotated"]
unrotated <- function(x) x[rotation == "unrotated"]
```

```{r check-rotation}
# Check that rotated and unrotated eofs
# predict the same fields (up to floating point errors)
bad_cells <- ceof[, predict(eof[[1]]), by = rotation] %>%
  .[, hgt := as.numeric(Re(hgt))] %>%
  dcast(time + lon + lat + lev ~ rotation, value.var = "hgt") %>%
  .[, dif := rotated - unrotated] %>%
  .[abs(rotated - unrotated)  > 1e-10]

if (nrow(bad_cells) != 0) {
  stop("Rotated and unrotated eofs are not equivalent!")
}

ceof <- ceof[rotation == "rotated"][, rotation := NULL][]
```

(ref:rotations-cap) a) Correlation of the Real and Imaginary cEOF1 with wave 1 amplitude of Total Ozone Column averaged between `r knitr::combine_words(LatLabel(o3wave_lats))` and b) $R^2$ between the Real and Imaginary cEOF2 and the ONI index for different values of rotation. The vertical line marks the selected rotation, which maximises de correlation.

```{r rotations, fig.cap="(ref:rotations-cap)", fig.height=2.5, fig.width=width_column, include=FALSE}
r2 <- expression(`$r^2$` = paste("", "r", phantom()^{
  paste("2")
}, "", ""))

rotations_o3 %>%
  ggplot(aes(rotation, cor)) +
  geom_vline(xintercept = best_rotation_cEOF1) +
  geom_hline(yintercept = 0,  size = 0.2, color = "gray50") +
  geom_line(aes(color = part)) +
  scale_x_continuous("Rotation", breaks = seq(-pi, pi, by = 90*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  scale_y_continuous("Correlation") +
  scale_color_brewer(NULL, palette = "Dark2", labels = lab_cplx) +

  rotations_cEOF2 %>%
  melt(id.vars = "rotation") %>%
  ggplot(aes(rotation, value^2)) +
  geom_hline(yintercept = 0,  size = 0.2, color = "gray50") +
  geom_vline(xintercept = best_rotation_cEOF2) +
  geom_line(aes(color = variable)) +
  scale_x_continuous("Rotation", breaks = seq(-pi, pi, by = 90*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  scale_y_continuous(r2) +
  scale_color_brewer(NULL, palette = "Dark2", labels = lab_cplx) +

  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "a", tag_suffix = ")")

```

While we compute these complex principal components using data from 1979 to 2019, we extended the complex time series back to the 1950 -- 1978 period by projecting monthly geopotential height zonal anomalies standardised by level south of 20ºS onto the corresponding spatial patterns.

We performed linear regressions to quantify the association between the cEOFs and other variables (e.g. geopotential height, temperature, precipitation, and others).
For each cEOF, we computed regression maps by fitting a multiple linear model involving both the real and the imaginary part.
To obtain the linear coefficients of a variable $X$ with the imaginary and real parts of each cEOF we fit the equation

$$
X(\lambda, \phi, t) = \alpha(\lambda, \phi) \operatorname{Im}(\mathrm{cEOF}) + \beta(\lambda, \phi) \mathrm{Re}(\mathrm{cEOF}) + X_0(\lambda, \phi) +  \epsilon(\lambda, \phi, t)
$$

where $\lambda$ and $\phi$ are the longitude and latitude, $t$ is the time, $\alpha$ and $\beta$ are the linear regression coefficients for imaginary and real parts respectively, $X_0$ and $\epsilon$ are the constant and error terms respectively.

We evaluated statistical significance using a two-sided t-test and, in the case of regression maps, p-values were adjusted by controlling for the False Discovery Rate [@benjamini1995; @wilks2016] to avoid misleading results from the high number of regressions [@walker1914; @katz1991].

## Computation procedures

We performed all analysis in this paper using the R programming language [@rcoreteam2020], using data.table [@dowle2020] and metR [@campitelli2020] packages.
All graphics are made using ggplot2 [@wickham2009].
We downloaded data from reanalysis using the ecmwfr package [@hufkens2020] and indices of ENSO and Indian Ocean Dipole with the rsoi package [@albers2020].
The paper was rendered using knitr and rmarkdown [@xie2015; @allaire2020].

# Results

## cEOF spatial patterns {#spatial}

```{r ceofs-1, fig.cap = "Spatial patterns for the two leading cEOFs of SON zonal anomalies of geopotential height at 50 hPa and 200 hPa for the 1979 -- 2019 period. The shading (contours) corresponds to real (imaginary) part. Arbitrary units.", fig.width=width_column, fig.height=width_column*1.2}

gdata <- ceof[, eof[[1]]$right, by = .(season)] %>% 
  copy() %>% 
  .[, hgt := hgt/sd(abs(hgt))]

var <- ceof[, eof[[1]]$sdev, by = .(season)] %>%
  .[, setNames(paste0(cEOF, " (", scales::percent(r2), ")"),
               cEOF)]

gdata %>%
  ggperiodic::periodic(lon = c(0, 360)) %>%
  ggplot(aes(lon, lat)) +

  geom_contour_fill(aes(z = Re(hgt), fill = ..level..),
                    breaks = AnchorBreaks(0, 1, exclude = 0)) +
  geom_contour2(aes(z = Im(hgt),  linetype = factor(sign(..level..))),
                breaks = AnchorBreaks(0, 1, exclude = 0)) +

  scale_fill_divergent_discretised(guide = guide_colorsteps_bottom()) +
  geom_qmap() +
  scale_x_longitude(labels = NULL) +
  scale_y_latitude(labels = NULL,
                   minor_breaks = NULL) +
  scale_linetype_manual(values = c("1" = 1, "-1" = 2),
                        labels = c("1" = "+", "-1" = "-"),
                        guide = "none") +
  facet_grid( lev  ~  cEOF,
              labeller = labeller(lev = lab_lev,
                                  cEOF = var)) +
  theme(legend.title = element_blank()) +
  geom_coords() +
  coord_polar() +

  tag_facets(tag = "cr")

```

```{r era_more}
pattern <- ceof$eof[[1]]$right %>%
  sep_ReIm() %>%
  setnames("hgt", "EOF")

era_more <- rbind( 
  ReadNetCDF(ERA5_BE(), vars = c(hgt = "z"),
             subset = list(latitude = range(pattern$lat),
                           level = list(200, 50))),
  ReadNetCDF(ERA5_geopotential(), vars = c(hgt = "z"),
             subset = list(latitude = range(pattern$lat),
                           level = list(200, 50)))) %>%
  normalise_coords() %>%
  .[season(time) == "SON"] %>%
  .[, w := season_weights(time)] %>% 
  .[, .(hgt = mean(hgt*w)), by = .(lev, lon, lat, time = seasonally(time))] %>%
  .[, season := season(time)] %>%
  .[, hgt := hgt/sd(hgt), by = .(lev)] %>%
  .[pattern, on = .NATURAL, allow.cartesian = TRUE] %>%
  .[, .(hgt = weighted.mean(hgt*EOF, w = cos(lat*pi/180))),  by = .(part, cEOF, time)]

```

(ref:extended-series-cap) Time series of the two leading cEOFs of SON zonal anomalies of geopotential height at 50 hPa and 200 hPa. cEOF1 (row a) and cEOF2 (row b) separated in their real (column 1) and imaginary (column 2) components. Dark straight line is the linear trend. Black horizontal and vertical line mark the mean value and range of each time series, respectively.

```{r extended-series, fig.cap = "(ref:extended-series-cap)", fig.height=3, fig.width = 6, fig.env = "figure*"}
# Check that the extended index coincides with the
# original index.
cors <- era_more %>%
  .[sep_ReIm(ceof$eof[[1]]$left), on = c("time", "cEOF", "part")] %>%
  na.omit() %>%
  .[, cor(hgt, i.hgt), by = .(cEOF, part)]

if (any(cors$V1 < 0.98)) {
  stop("Extended indices didn't work!")
}

trends <- era_more %>%
  .[, FitLm(hgt, time, se = TRUE), by = .(part, cEOF)] %>%
  rm_intercept() %>%
  .[, p_val := Pvaluate(estimate, std.error, df)] %>%
  .[, p_report := report_p(p_val)]

# era_less <-   ceof %>%
#   .[, eof[[1]]$left, by = season] %>%
#   sep_ReIm(hgt) %>%
#   .[season == "SON"] %>% 
#   .[, hgt := as.numeric(scale(hgt)), by = .(part, cEOF)]

era_more %>%
  copy() %>% 
  .[, ref := year(time) > 1979] %>% 
  # .[, hgt := scale(hgt, center = FALSE, scale = sd(hgt[ref])), by = .(part, cEOF)] %>%
  ggplot(aes(time, hgt*1e3)) +
  # Zero line + rangeframe
  geom_hline(data = ~.x[, mean(hgt)*1e3, by  = .(cEOF, part)], 
             aes(yintercept = V1), size = 0.2, color = "gray30") +
  ggthemes::geom_rangeframe(sides = "r", size = 0.2, color = "gray30") +
  geom_line(aes(color = part)) +
  geom_point(aes(color = part), size = 0.25) +
  # geom_line(data = era_less) +
  scale_alpha_manual(values = c(1, 0.2)) +
  geom_smooth(method = "lm", aes(color = stage(part, after_scale = scales::muted(color))),
              show.legend = FALSE, size = 0.8) +
  scale_x_date(NULL, date_labels = "%Y",
               breaks = seq(as.Date("1950-01-01"),
                            as.Date("2020-01-01"), by = "10 years"),
               minor_breaks = seq(as.Date("1950-01-01"),
                                  as.Date("2020-01-01"), by = "2 years")) +
  scale_y_continuous(NULL, minor_breaks = NULL, sec.axis = dup_axis()) +
  scale_color_brewer(NULL, palette = "Dark2", 
                     guide = guide_legend(override.aes = list(fill = NA))) +
  facet_grid(cEOF ~ part) +
  theme(panel.spacing = grid::unit(1.5, "lines"),
        strip.placement = "outside", 
        panel.grid.minor.x = element_line(linetype = 2))  +
  tag_facets("rc") +
  panel_background 
```

To describe the variability of the circulation zonal anomalies, the spatial and temporal parts of the first two leading cEOFs of zonal anomalies of geopotential height at 50 hPa and 200 hPa, computed jointly at both levels, are shown in Figures \@ref(fig:ceofs-1) and \@ref(fig:extended-series).
The first mode (cEOF1) explains 82% of the variance of the zonally anomalous fields, while the second mode (cEOF2) explains a smaller fraction (7%).
In the spatial patterns (Fig. \@ref(fig:ceofs-1)), the real and the imaginary components are in quadrature by construction, so that each cEOF describe a single wave-like pattern whose amplitude and position (i.e. phase) is controlled by the magnitude and phase of the temporal cEOF.
The wave patterns described by these cEOFs match the patterns seen in the traditional EOFs of Figure \@ref(fig:eof-naive).

The cEOF1 (Fig. \@ref(fig:ceofs-1) column a) is a hemispheric wave 1 pattern with maximum amplitude at high latitudes.
At 50 hPa the Real cEOF1 has the maximum of the wave 1 at 150ºE and at 200 hPa, the maximum is located at around 175ºE indicating a westerly shift in phase.
The cEOF2 (Fig. \@ref(fig:eof-naive) column b) shows also a zonal wave-like structure with maximum amplitude at high latitudes, but with shorter spatial scales.
In particular, the dominant structure at both levels is a wave 3 but with larger amplitude in the pacific sector.
This modulated amplitude more evident at 200 hPa.
There is no apparent phase shift with height but the amplitude of the pattern is greatly reduced in the stratosphere, suggesting that this barotropic mode represents mainly tropospheric variability.

There is no significant simultaneous correlation between cEOFs time series.
Both cEOFs show year-to-year variability but show no evidence of decadal variability (Fig. \@ref(fig:extended-series)).
Because the geopotential fields that enter into the cEOFs algorithm are anomalies with respect to the zonal mean instead of the time mean, the cEOFs have non zero temporal mean.
However, cEOF2's mean is almost zero, which indicates that only cEOF1 includes variability that significantly projects onto the mean zonally anomalous field.

A significant upward trend in the real component of cEOF1 is evident (Fig. \@ref(fig:extended-series)a.1, p-value `r trends[part == "Real" & cEOF == "cEOF1"]$p_report`) while there is no significant trend in any of the complex components of cEOF2.
The positive trend in the Real cEOF1 translates into a positive trend in cEOF1 magnitude, but not systematic change in phase (not shown).
This long-term change indicates an increase in the magnitude of the high latitude zonal wave 1.
In agreement, @raphael2003 detected a step change after around 1975 in the temporal evolution of leading EOF computed from August-September-October 500 hPa zonal geopotential height anomalies, which is similar to the cEOF1.

## cEOFs Regression maps {#regressions}

In the previous section, cEOFs were applied to zonal anomalies, that is anomalies derived by removing the zonally mean values in order to isolate the main characteristics of the main zonal waves characterizing the circulation in the SH.
In this section, regression fields were computed using the full fields of the variables in order to describe the influence of the cEOFs on the temporal anomalies.

```{r compute-regression}

era5 <- ERA5_temperature() %>% 
  ReadNetCDF(vars = c("t"), 
             subset = list(level = list(50, 200))) %>% 
  normalise_coords() %>% 
  .[season(time) == "SON"] %>%
  .[, w := season_weights(time)] %>% 
  .[, .(t = mean(t*w)), by = .(lev, lon, lat, time = seasonally(time))] %>% 
  .[era5, on = .NATURAL]

regr <- ceof[, eof[[1]]$left, by = .(season)] %>%
  .[season == "SON"] %>%
  sep_ReIm(format = "wide") %>%
  .[, hgt := NULL] %>%
  .[, `:=`(Real = as.numeric(scale(Real)), Imaginary = as.numeric(scale(Imaginary))),
    by = .(season, cEOF)] %>%
  .[melt(era5, id.vars = c("time", "lon", "lat", "lev")), on = "time", allow.cartesian = TRUE] %>%
  .[, FitLm(value, Imaginary, Real, se = TRUE),
    by = .(lev, lon, lat, variable, cEOF)] %>%
  rm_intercept() %>%
  .[, term := factor_ReIm(term)]
```

```{r plot_regr}
plot_regr <- function(data, which_cEOF, which_levs, which_var) {
  data <- data %>%
    .[ cEOF %in% which_cEOF] %>%
    .[variable == which_var]

  var_title <- switch(which_var,
                      hgt = "Geopotential height",
                      t = "Temperature")

  if (length(which_cEOF) > 1) {
    facet <- ggh4x::facet_nested(cEOF ~  term)
  } else {
    facet <- ggh4x::facet_nested(lev ~ cEOF +  term, 
                                 labeller = labeller(lev = lab_lev))
  }

  if ("lev" %in% colnames(data)) {
    data <- data[lev %in% which_levs]
  } else {
    data[, lev := which_levs]
  }

  data[lat <= -20] %>%
    # .[term == "I" & lev == 50] %>%
    .[, p.value := Pvaluate(estimate, std.error, df, "fdr"), by = .(lev, cEOF, term)] %>%
    periodic(lon = c(0, 360)) %>%
    ggplot(aes(lon, lat)) +
    geom_contour_fill(aes(z = estimate), breaks = AnchorBreaks(exclude = 0),
                      global.breaks = FALSE) +
    geom_contour_tanaka(aes(z = estimate), 
                        breaks = AnchorBreaks(exclude = 0), global.breaks = FALSE) +
    scale_fill_divergent(var_title, breaks = AnchorBreaks(0, exclude = 0),
                         guide = guide_colorstrip_bottom(15)) +
    geom_contour_pval(aes(z = p.value)) +

    geom_qmap() +

    scale_x_longitude(labels = NULL) +
    scale_y_latitude(labels = NULL,
                     minor_breaks = NULL) +
    grid_y +
    facet +
    geom_coords() +
    coord_polar() +
    tag_facets("rc")
}

mgp <- latex2exp::TeX('Geopotential height ($m^2s^{-1}$)')
```

(ref:eof1-regr-gh-cap) Regression of SON geopotential height anomalies ($m^2s^{-1}$) with the (column 1) real and (column 2) imaginary parts of the first cEOF for the 1979 -- 2019 period at (row a) 50 hPa and (row b) 200 hPa. These coefficients come from multiple linear regression involving the real and imaginary parts. Areas marked with dots have p-values smaller than 0.01 adjusted for False Detection Rate.

```{r eof1-regr-gh, fig.cap = "(ref:eof1-regr-gh-cap)", fig.height=1.2*width_column, fig.width=width_column, dependson="plot_regr"}
plot_regr(regr , "cEOF1", c(200, 50), "hgt") +
  scale_fill_divergent(mgp,
                       breaks = AnchorBreaks(0, exclude = 0),
                       guide = guide_colorsteps_bottom(15, even.steps = TRUE), 
                       limits = c(-2000, 2000)) 
```

Figure \@ref(fig:eof1-regr-gh) shows regression maps of SON geopotential height anomalies upon cEOF1.
At 50 hPa (Figure \@ref(fig:eof1-regr-gh) row a), the Real cEOF1 is associated with a monopole centred over the Ross Sea.
The Imaginary cEOF1 is associated with a clear wave 1 pattern with maximum over the coast of East Antarctica.

At 200 hPa (Figure \@ref(fig:eof1-regr-gh) row b) the Real cEOF1 shows a single area of positive anomalies spanning West Antarctica surrounded by opposite anomalies in lower latitudes with is centre shifted slightly eastward compared with the upper level anomalies.
The Imaginary cEOF2.
The Imaginary cEOF1 shows a much more zonally symmetrical pattern resembling the negative SAM phase [@fogt2020].
Both components of cEOF1 could be associated to zonal anomalies in the upper troposphere jet stream intensity.

Therefore, the magnitude and phase of the cEOF1 are associated with the magnitude and phase of a zonal wave only in the stratosphere.
While in the troposphere, they are associated with slightly off-centre monopoles.

(ref:eof2-regr-gh-cap) Same as Figure \@ref(fig:eof1-regr-gh) but for cEOF2.

```{r eof2-regr-gh, fig.cap = "(ref:eof2-regr-gh-cap)", fig.height=1.3*width_column, fig.width=width_column, dependson="plot_regr"}
plot_regr(regr, "cEOF2", c(200, 50), "hgt") +
  scale_fill_divergent(mgp, 
                       breaks = AnchorBreaks(0, 200, exclude = 0),
                       guide = guide_colorsteps_bottom(15, even.steps = TRUE), 
                       limits = c(-1000, 1000))
```

Figure \@ref(fig:eof2-regr-gh) shows the regression maps of geopotential height anomalies upon the cEOF2.
Wave trains similar to those identified for cEOF2 patterns (Fig \@ref(fig:ceofs-1)) can be distinguished.
Regressed anomalies associated with the Real cEOF2 are 90º out of phase with those associated with the Imaginary cEOF2.
All fields have a dominant zonal wave 3 limited to the western hemisphere, over the Pacific and Atlantic Oceans.
cEOF2 then represents an equivalent barotropic wave train that is very similar to the the PSA Patterns [@mo2001].
Comparing the location of the positive anomaly near 90ºW in column b of Figure \@ref(fig:eof2-regr-gh) with Figures 1.a and b from @mo2001, the Real cEOF2 regression map could be identified with PSA2, while the Imaginary cEOF2 resembles PSA1.

## cEOFs relationship with known modes of variability {#other-variables}

### SAM

(ref:sam-eof-vertical-cap) Coefficient of determination between the real and imaginary part of each cEOF and the SAM, Asymmetric SAM (A-SAM) and Symmetric SAM (S-SAM) indices computed at each level according to \@@campitelli2022 for the 1979 -- 2019 period. Thick lines represent estimates with p-value \< 0.01 corrected for False Detection Rate [@benjamini1995].

```{r sam-eof-vertical, fig.cap = "(ref:sam-eof-vertical-cap)", fig.height=4, fig.width=width_column}
time <- ceof[, eof[[1]]$left, by = .(season)] %>%
  sep_ReIm() %>%
  .[, hgt := scale(hgt), by = .(cEOF, part)]

sam_r2 <- sams %>%
  .[time, on = .NATURAL, allow.cartesian = TRUE] %>%
  na.omit() %>% 
  .[, correlate(hgt, estimate), by = .(lev, term, part, cEOF)]

max_r2 <- sam_r2[part == "Imaginary" & term == "asym"] %>% 
  .[, .SD[which.max(estimate^2)], by = .(cEOF)]

sam_r2 %>%
  .[, p.value_a := p.adjust(p.value, "fdr")] %>% 
  ggplot(aes(lev, estimate^2)) +
  geom_vline(xintercept = c(50, 200),  size = 0.2, color = "gray50") +
  # geom_point(data = ~.x[p.adjust(p.value, "fdr") < 0.01], aes(color = term)) +
  geom_line(data = ~copy(.x)[p.value_a > 0.01, estimate := NA], 
            aes(colour = stage(term, after_scale = scales::alpha(colour, 0.5))), 
            size = 2) +
  geom_line(aes(color = term, group = term)) +

  scale_x_level("Level (hPa)", minor_breaks = NULL, 
                sec.axis = sa_height_axis(breaks = seq(0, 50, by = 10),
                                          labels = function(x) paste0("~", x, " km"))) +
  scale_y_continuous(r2, limits = c(0, 1), 
                     guide = guide_axis(check.overlap = TRUE)) +
  scale_size_manual("p-value", values = c("TRUE" = 1, 
                                          "FALSE" = 0.5), 
                    guide = "none") +
  scale_color_brewer(NULL, palette = "Dark2", labels = c(full = "SAM",
                                                         asym = "A-SAM",
                                                         sym = "S-SAM")) +
  facet_grid(part ~ cEOF) +
  coord_flip() +
  tag_facets("rc") +
  panel_background +
  theme(panel.spacing = grid::unit(1, "lines"), 
        strip.placement = "outside")
```

To explore the relationship between SAM and the modes described from the cEOF, we compute the coefficient of determination between the cEOFs timeseries and the three SAM indices (SAM, A-SAM and S-SAM) at each vertical level (Fig. \@ref(fig:sam-eof-vertical)).
The SAM index is statistically signficant correlated with the Real cEOF1 in all levels, and with the Imaginary cEOF1 and Imaginary cEOF2 in the troposphere.
Correlations between SAM and the Real cEOF2 are non-significant.

The relationship between the tropospheric SAM and cEOF1 is explained entirely by the zonally symmetric component of the SAM as shown by the low and statistically non-significant correlations between the A-SAM and either the Real or Imaginary cEOF1 and the high correlation with the S-SAM below 100 hPa.
In the stratosphere, the Real cEOF1 is correlated with both A-SAM and S-SAM, while the Imaginary cEOF1 is highly correlated only with the A-SAM.
These correlations are consistent with the regression patterns of geopotential height in Figure \@ref(fig:eof1-regr-gh) and their comparison with those obtained for SAM, A-SAM and S-SAM by \@@campitelli2022.

```{r eof-filtered, eval=FALSE, fig.cap="Spatial pattern of the leading EOF of 200 hPa geopotential height with the variability of cEOF2 filtered out South of 20º. Arbitrary units.", fig.height=2.5, fig.width=2.5, include=FALSE}
cEOF2 <- ceof$eof[[1]]$left[cEOF == "cEOF2"] %>% 
  copy() %>% 
  .[, cEOF := NULL] %>% 
  .[]

eof_filtered <- era5 %>% 
  .[sep_ReIm(cEOF2, format = "wide")[, hgt := NULL], on = .NATURAL] %>% 
  .[lev == 200] %>% 
  .[lat <= -20] %>% 
  .[season(time) == "SON"] %>% 
  .[, hgt := resid(.lm.fit(cbind(1, Real, Imaginary), hgt)), .(lon, lat, lev, season(time))] %>%
  .[, eof := Anomaly(hgt)*sqrt(cos(lat*pi/180)), by = .(lon, lat, lev, season(time))] %>% 
  EOF(eof ~ lon + lat | time, n = 1, data = .)

eof_filtered_cor <- eof_filtered$right[sams, on = "time"] %>% 
  .[lev == 200] %>% 
  .[, correlate(estimate, eof), by = term]

eof_filtered %>%
  .$left %>%
  periodic(lon = c(0, 360)) %>%
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = eof), breaks = AnchorBreaks(0, exclude = 0)) +
  geom_contour_tanaka(aes(z = eof), breaks = AnchorBreaks(0, exclude = 0)) +
  scale_fill_divergent(guide = "none") +
  geom_qmap() +
  scale_x_longitude() +
  scale_y_latitude() +
  geom_coords() +
  coord_polar() +

  theme(axis.text = element_blank())

eof_filtered_cor %>% 
  .[, .(term, text)] %>%
  .[, term := factor(term, labels = c("SAM", "A-SAM", "S-SAM"))] %>%
  knitr::kable(col.names = c("Índice SAM", "Correlación"))
```

In the case of cEOF2, the moderate correlation between SAM and Imaginary cEOF2 becomes extremely high only when the zonally asymmetric variability of the SAM is considered.
The Imaginary cEOF2 explains up to `r scales::percent(max_r2[cEOF == "cEOF2"]$estimate^2, 2)` of A-SAM variance, reached the maximum value at `r max_r2[cEOF == "cEOF2"]$lev` hPa (Figure \@ref(fig:sam-eof-vertical).b2).
Such extremely high correlation between the asymmetric SAM and the Imaginary cEOF2 suggests that these might be different ways of characterising the same phenomenon.

### PSA

(ref:psa-eof2-cap) Correlations between the Real and Imaginary parts of cEOF2 and the PSA1 and PSA2 modes computed as the second and third EOFs of seasonal geopotential height anomalies [following @mo2001] for the 1979 -- 2019 period. 95% confidence intervals in parenthesis. p-values lower than 0.01 in bold.

```{r psa-eof2, fig.cap = "(ref:psa-eof2-cap)", fig.width=7, fig.height=3, fig.env="figure*"}
psa2 <- psa$left[PC == "PSA2"][season(time) == "SON"]
cEOF2 <- ceof$eof[[1]]$left[cEOF == "cEOF2"] %>% 
  copy() %>% 
  .[, cEOF := NULL] %>% 
  .[]

rotations_psa <- lapply(angles, function(a) {
  cEOF2 %>% 
    copy() %>% 
    .[, hgt := rotate(hgt, a)] %>% 
    .[psa2, on = "time"] %>% 
    .[, cor(value, Re(hgt))] 
})

psa_angle <- angles[which.min(unlist(rotations_psa))]

make_numeric <- function(x) {
  x <- strsplit(x, "\ ")
  vapply(x, function(x) as.numeric(x[[1]]), numeric(1))
}

background_palette <- grDevices::colorRampPalette(c(scales::muted("blue"),
                                                    "white",
                                                    scales::muted("red")),
                                                  space = "Lab")

table <- cEOF2[psa_time, on = "time"] %>% 
  .[PC != "SAM"] %>% 
  copy() %>% 
  sep_ReIm() %>% 
  .[, correlate(value, hgt), by = .(PC, part)]

p.val <- table %>% 
  dcast(PC ~ part, value.var = "p.value")

table %>% 
  dcast(PC ~ part, value.var = "text") %>% 
  kbl2(booktabs = TRUE, caption = "(ref:psa-eof2-cap)") %>%
  column_spec2(2:3, bold = p.val[[.col]] < 0.01) %>%
  add_header_above(c("", "cEOF2" = 2)) %>%
  kable_classic()

# cEOF2[psa_time, on = "time"] %>% 
#   .[PC != "SAM"] %>% 
#   copy() %>% 
#   # .[, hgt := rotate(hgt, psa_angle)] %>%
#   sep_ReIm() %>% 
#   ggplot(aes(value, hgt)) +
#   geom_hline(yintercept = 0, size = 0.2, color = "gray50") +
#   geom_vline(xintercept = 0, size = 0.2, color = "gray50") +
#   geom_point() +
#   geom_smooth(method = "lm") +
#   geom_smooth(method = "lm") + 
#   scale_y_continuous("EOF2") +
#   scale_x_continuous("PSA") +
#   facet_grid(part ~ PC) +
#   panel_background +
#   tag_facets("rc")
```

Due to the similarity between the PSA patterns and the cEOF2 spatial patterns (Fig. \@ref(fig:eof2-regr-gh)), we study the relationship between these modes.
Table \@ref(tab:psa-eof2) shows the correlations between the two PSA indices and the timeseries for Real and Imaginary phase of cEOF2.
As anticipated by Figure \@ref(fig:eof2-regr-gh), there is a strong correlation between PSA1 and Imaginary cEOF2, and between PSA2 and Real cEOF2.
Conversely, there is no relationship between PSA1 and Real cEOF2, and between PSA2 and Imaginary cEOF2.
As a result, cEOF2 represents well both the spatial structure and temporal evolution of the PSA modes, as well as it is possible to make a rather clean association between its two phases and the two PSA modes.
So, It could be concluded that the same particular rotation of cEOF2 that maximises the association between cEOF2 parts and PSA modes, is the one which maximises the relationship between ENSO and Imaginary cEOF2.

The reason the conventional EOF analysis arrives at the same separation than our particular cEOF rotation is probably the fact that not all phases are equally frequent.
To visualize that, Figure \@ref(fig:phase-histogram) shows an histogram that counts the number of SON years in which the cEOF2 was close to each of the four particular phases (positive/negative Real/Imaginary), with the observations for each season marked as rugs on the horizontal axis.
For instance, years with cEOF2 phase within 45º of 0º are nearest the "positive Real" phase.
About two thirds of time cEOF2 has a phase similar to either the negative or positive Imaginary phase.
It could be seen that Imaginary phase is the most common phase and It is also, by construction, the direction that has the maximum relationship with ENSO.
Therefore, the Imaginary cEOF2 explains more variance than the Real cEOF2 and conventional EOF analysis will tend to separate the two.

(ref:phase-histogram-cap) Histogram of phase distribution of cEOF2 for the 1979 -- 2019 period. Bins are centred at 90º, 0º, -90º, -180º with a binwidth of 90º. The small vertical lines near the horizontal axis mark the observations.

```{r phase-histogram, fig.cap = "(ref:phase-histogram-cap)", fig.width=width_column*.8, fig.height=2.5}
N <- nrow(cut(ceof$eof[[1]], 2)$left)
cut(ceof$eof[[1]], 2)$left %>% 
  .[, arg := Arg(hgt)] %>% 
  .[, w := Mod(hgt)/mean(Mod(hgt))] %>% 
  qwrap(arg = c(-pi, pi) ~  c(-pi - 1/4*pi, pi - 1/4*pi)) %>% 
  ggplot(aes(arg)) +
  geom_histogram(binwidth = 90*pi/180, center = 0, size = 0.2,
                 color = "black", fill = "gray70") +
  annotate("text",
           label = c("- Real", "- Imaginary", "+ Real", "+ Imaginary"),
           color = "black", size = 3,
           y = 0.5, x = c(-pi, -pi/2, 0, pi/2), angle = 90, hjust = 0) +
  geom_rug() +
  scale_y_continuous("Number of years", 
                     sec.axis = sec_axis(~.x/N, 
                                         name = "Proportion of years", 
                                         labels = scales::percent)) +
  scale_x_continuous("Phase", trans = scales::reverse_trans(),
                     breaks = seq(-pi - 1/4*pi, pi - 1/4*pi, by = 45*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  no_grid
```

The result obtained here is similar to the one obtained by @irving2016.
They show that the phase distribution of the PSA-like variability obtained by them is bimodal and isolate the PSA pattern from the rest of the PSA-like variability by selecting events that are near the peaks of the distribution (compare our Figure \@ref(fig:phase-histogram) with their Figure 6).

The advantage of our method is that it is much simpler to implement, it provides magnitude and phase naturally, and it facilitates the description of this mode as a propagating wave instead of as standing oscillations.
As a consequence, the cEOF2 offers an alternative way of representing the PSA which has several advantages over using the second and third EOFs.

### Temperature and Ozone

(ref:eof1-regr-t-cap) Same as Figure \@ref(fig:eof1-regr-gh) but for air temperature (K).

```{r eof1-regr-t, fig.cap = "(ref:eof1-regr-t-cap)",  dependson="plot_regr", fig.height=1.3*width_column, fig.width=width_column}
plot_regr(regr, "cEOF1", c(200, 50), "t") +
  scale_fill_divergent("Temperature (K)", breaks = AnchorBreaks(0, exclude = 0),
                       guide = guide_colorsteps_bottom(15, even.steps = TRUE), 
                       limits = c(-4.5, 4.5))
```

```{r t}
t <- ERA5_temperature() %>% 
  ReadNetCDF(vars = c("t"), 
             subset = list(latitude = o3wave_lats)) %>% 
  normalise_coords() %>% 
  .[season(time) == "SON"] %>% 
  .[, .(t = weighted.mean(t, cos(lat*pi/180))), by = .(lon, lev, time)]

o3_temp <- ERA5_ozone() %>% 
  ReadNetCDF(vars = c("o3"), 
             subset = list(latitude = o3wave_lats)) %>% 
  normalise_coords() %>% 
  .[season(time) == "SON"] %>% 
  .[, .(o3 = weighted.mean(o3, cos(lat*pi/180))), by = .(lon, lev, time)]

t <- t[o3_temp, on = .NATURAL]
rm(o3_temp)

t <- t %>% 
  .[, w := season_weights(time)] %>% 
  .[, .(t   = mean(t*w),
        o3  = mean(o3*w)), by = .(lev, lon, time = seasonally(time))]
```

(ref:t-vertical-cap) Regression of SON anomalies of mean air temperature (shaded, Kelvin) and ozone mixing ratio (contours, negative contours with dashed lines, labels in parts per billion by mass) averaged between `r knitr::combine_words(LatLabel(o3wave_lats))` with the (a) real and (b) imaginary parts of the first cEOF for the 1979 -- 2019 period.

```{r lable_placement_min}
label_placement_min <- function(direction = c("vertical", "horizontal"))  {
  direction <- direction[1]
  function(x, y) {
    browser()
    y <- x[[1]]$y
    x <- x[[1]]$x

    if (direction == "vertical") {
      return(which.min(y))
    }
    else if (direction == "horizontal") {
      return(which.min(x))
    }
  }
}
```

```{r t-vertical, fig.cap = "(ref:t-vertical-cap)", fig.height=2.5, fig.width=width_column}
t %>% 
  melt(id.vars = c("lev", "time", "lon")) %>% 
  .[, value := Anomaly(value), by = .(time, lev, variable)] %>%
  .[cut(ceof$eof[[1]], 1)$left, on = "time"] %>% 
  sep_ReIm(format = "wider") %>% 
  .[, ":="(Real = as.numeric(scale(Real)), Imaginary = as.numeric(scale(Imaginary))), by = .(cEOF)] %>% 
  .[, FitLm(value, Real, Imaginary, se = TRUE), by = .(lev, lon, variable)] %>% 
  rm_intercept() %>% 
  .[, term := factor_ReIm(term)] %>% 
  ggplot(aes(lon, lev)) +
  geom_contour_fill(data = ~.x[variable == "t"],
                    aes(z = estimate, fill = ..level..),
                    breaks = AnchorBreaks(0, binwidth = 0.25, exclude = 0, sym = TRUE)) +
  geom_contour_tanaka(data = ~.x[variable == "t"],
                      aes(z = estimate),
                      breaks = AnchorBreaks(0, binwidth = 0.25, exclude = 0, sym = TRUE)) +
  geom_contour2(data = ~.x[variable == "o3"],
                aes(z = estimate, linetype = factor(-sign(..level..))),
                breaks = AnchorBreaks(0, exclude = 0)) +
  geom_text_contour(data = ~.x[variable == "o3"],
                    aes(z = estimate*1e9, label = signif(..level.., 2)),
                    size = 2,
                    skip = 1, stroke = 0.15, stroke.colour = "white",
                    # label.placement = label_placement_min(),
                    breaks = AnchorBreaks(0, exclude = 0)) +

  scale_fill_divergent_discretised("Air temperature (K)",
                                   # limits = c(-2.25, 2.25),
                                   labels = one_less,
                                   guide = guide_colorsteps_bottom(15, show.limits = TRUE)) +
  scale_y_level(minor_breaks= NULL,
                sec.axis = sa_height_axis(breaks = seq(0, 50, by = 10),
                                          labels = function(x) paste0("~", x, " km"))) +
  scale_linetype(guide = "none") +
  scale_x_longitude(ticks = 60) +
  facet_grid(~term) +
  axis_labs_smol +
  coord_cartesian(clip = "off") +
  tag_facets() +
  NULL
```

The relation between cEOFs and air temperature was also evaluated.
Figure \@ref(fig:eof1-regr-t) shows regression patterns of air temperature at 50hPa and 200hPa onto cEOF1.
In both levels, the Real cEOF1 is associated with a positive monopole over the South Pole with its centre moved slightly towards 150ºE (Fig. \@ref(fig:eof1-regr-t) column 1).
On the other hand, the regression maps on the Imaginary cEOF1 show a more clear wave 1 pattern with its maximum around 60ºE.
The distribution of temperature regression coefficients at 50 hPa and at 200 hPa mirror the geopotential height regression maps at 50 hPa (Fig. \@ref(fig:eof1-regr-gh)).

Figure \@ref(fig:t-vertical) shows the vertical distribution of the regression coefficients on cEOF1 from zonal anomalies of air temperature and zonal anomalies of ozone mixing ratio averaged between `r knitr::combine_words(LatLabel(o3wave_lats))`.
Temperature zonal anomalies associated with cEOF1 show a clear wave 1 pattern for both real and imaginary components throughout the atmosphere above 250 hPa with a change in sign above 10 hPa.
Following hydrostatic balance, this is the level in which the geopotential anomaly have maximum amplitude (not shown).

The maximum ozone anomalies are co-located with the minimum temperature anomalies above 10 hPa and with the maximum temperature anomalies below 10 hPa (Fig. \@ref(fig:t-vertical)).
Therefore, the ozone zonal wave 1 is anticorrelated with the temperature zonal wave 1 in the upper stratosphere, and directly correlated in the upper stratosphere.
This change in phase is observed in ozone anomalies forced by planetary waves that reach the stratosphere.
In the photochemically-dominated upper stratosphere, cold temperatures inhibit the destruction of ozone.
On the other hand, in the advectively-dominated lower stratosphere, ozone anomalies are 90º out of phase with horizontal and vertical transport, which are in adittion 90º out of phase with temperature anomalies [@hartmann1979; @wirth1993; @smith1995].

```{r o3-t-phase}
# t %>% 
#   .[lat %between% c(-80, -35)] %>% 
#   .[, .(o3 = weighted.mean(o3, cos(lat*pi/180)),
#         t = weighted.mean(t, cos(lat*pi/180))), by = .(lev, time, lon)] %>%  
#   melt(id.vars = c("lev", "time", "lon")) %>% 
#   # .[, value := Anomaly(value), by = .(time, lev, variable)] %>%
#   .[cut(ceof$eof[[1]], 1)$left, on = "time"] %>% 
#   sep_ReIm(format = "wider") %>% 
#   .[, ":="(Real = as.numeric(scale(Real)), Imaginary = as.numeric(scale(Imaginary))), by = .(cEOF)] %>% 
#   .[, FitLm(value, Real, Imaginary, se = TRUE), by = .(lev, lon, variable)] %>% 
#   rm_intercept() %>% 
#   .[, term := factor_ReIm(term)] %>% 
#   .[, FitWave(estimate, 1), by = .(lev, variable, term)] %>% 
#   dcast(lev + term ~ variable, value.var = c("phase", "amplitude")) %>%
#   ggplot(aes(lev, acos(cos(phase_t - phase_o3)))) +
#   geom_line(aes(color = term)) +
#   scale_x_level() +
#   scale_colour_brewer(NULL, palette = "Dark2") +
#   scale_y_continuous("Wave-1 phase difference\nbetween Temperature and Ozone",
#                      breaks = seq(-pi, pi, by = 30*pi/180),
#                      labels = function(x) paste0(round(x*180/pi, 5), "º"))  +
#   coord_flip()
```

```{r o3_regr}
temp <- ceof$eof[[1]]$left %>% 
  .[cEOF == "cEOF1"] %>% 
  copy() %>% 
  sep_ReIm(format = "wider") %>% 
  .[, `:=`(Real = as.numeric(scale(Real)), 
           Imaginary = as.numeric(scale(Imaginary))),
    by = .(cEOF)]
o3_mean <- o3 %>% 
  .[, .(toc = mean(toc)), by = .(lon, lat)] %>% 
  .[, toc_z := Anomaly(toc), by = .(lat)]
o3_regr <- temp %>% 
  o3[., on = "time", allow.cartesian = TRUE] %>% 
  .[, FitLm(toc, Real, Imaginary, se = TRUE), by = .(lon, lat, cEOF)] %>% 
  rm_intercept() %>% 
  .[, term := factor_ReIm(term)]
```

(ref:o3-regr-cap) Regression of SON mean Total Ozone Column anomalies (shaded, Dobson Units) with the (a) real and (b) imaginary parts of the cEOF1 for the 1979 -- 2019 period. On contours, the mean zonal anomaly of Total Ozone Column (negative contours in dashed lines, Dobson Units). Areas marked with dots have p-values smaller than 0.01 adjusted for False Detection Rate.

```{r o3-regr, fig.cap = "(ref:o3-regr-cap)", fig.height=3, fig.width = width_column}
o3_regr %>% 
  periodic(lon = c(0, 360)) %>% 
  .[, p.value := Pvaluate(estimate, std.error, df, "fdr"), by = .(cEOF, term)] %>% 
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate, fill = ..level..),
                    breaks = AnchorBreaks(0, exclude = 0)) +
  geom_contour_tanaka(aes(z = estimate), breaks = AnchorBreaks(0, exclude = 0)) +
  geom_contour2(data = periodic(o3_mean, lon = c(0, 360)),
                aes(z = toc_z, linetype = factor(-sign(..level..))), 
                breaks = AnchorBreaks(0, exclude = 0), 
                bins = 5) +

  geom_contour_pval(aes(z = p.value)) +
  annotate("segment", 
           y  = o3wave_lats,
           yend = o3wave_lats,
           x = 0,
           xend = 360,
           size = 0.3, alpha = 0.8
  ) +
  geom_qmap() +
  scale_fill_divergent_discretised("Total Ozone Column (DU)", 
                                   limits = c(-30, 30), 
                                   breaks = AnchorBreaks(exclude = 0, binwidth = 5)(c(-30, 30)),
                                   guide = guide_colorsteps_bottom(15)) + 
  scale_x_longitude(labels = NULL) +
  scale_y_latitude(labels = NULL,
                   minor_breaks = NULL) +
  scale_linetype(guide = "none") +
  geom_coords() +
  coord_polar() +
  facet_grid(cEOF~term) +
  tag_facets("panel")

```

The regression maps of cEOF1 with fields of Total Ozone Column (TOC) (Fig.  \@ref(fig:o3-regr)) show zonal wave 1 pattens in TOC associated with both phases of cEOF1.
Climatologically, the springtime Ozone minimum is located off the South Pole and towards the Weddell Sea [@grytsai2011].
Thus, the Real cEOF1 regression pattern (Figure \@ref(fig:o3-regr)a) coincides with the climatological position of the ozone hole while the one for the Imaginary cEOF1 is shifted by 90º.

(ref:wave1-o3-cap) Relationship between cEOF1 amplitude and phase with amplitude and phase of the zonal wave 1 in Total Ozone Column averaged between `r knitr::combine_words(LatLabel(o3wave_lats))` for the period 1979 -- 2019.

```{r wave1-o3, fig.cap = "(ref:wave1-o3-cap)", fig.width=width_column, fig.height=width_column*.7, include = FALSE}
wave1_o3 <- o3 %>% 
  .[lat %between% o3wave_lats] %>% 
  .[, .(toc = mean(toc, na.rm = TRUE)), by = .(time, lon)] %>% 
  .[, FitWave(toc, 1), by = .(time)]

temp <- ceof$eof[[1]]$left %>% 
  .[cEOF == "cEOF1"] %>% 
  .[wave1_o3, on = "time"]

# temp %>% 
#   copy() %>% 
#   na.omit() %>% 
#   sep_ReIm() %>% 
#   .[, hgt := scale(hgt), by  = .(part)] %>% 
#   ggplot(aes(amplitude, hgt)) +
#   geom_point() +
#   geom_smooth(method = "lm", color = "black") +
#   facet_grid(cEOF ~ part) +
#   panel_background
wave1_cor <- rbind(temp %>%
                     .[, correlate(amplitude, Mod(hgt))] %>% 
                     .[, with := "amplitude"],

                   temp %>%
                     .[, correlate(phase, Arg(hgt))] %>% 
                     .[, with := "phase"])

wave_correlation <- temp %>% 
  na.omit() %>% 
  .[, .(cor = mean(cos(phase - Arg(hgt))),
        n = .N)] %>% 
  .[, correlation_statistics(cor, n)]

temp %>% 
  ggplot(aes(Mod(hgt), amplitude)) +
  scale_x_continuous("EOF1 Amplitude") +
  scale_y_continuous("O3 Wave-1\nAmplitude") +

  temp %>% 
  copy() %>%
  na.omit() %>% 
  qwrap(phase = c(0, 2*pi) ~ c(-pi, pi)) %>% 
  ggplot(aes(Arg(hgt), phase)) +
  scale_x_continuous("EOF1 Phase",
                     breaks = seq(-pi, pi, by = 15*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  scale_y_continuous("O3 Wave-1\nPhase",
                     breaks = seq(-pi, pi, by = 30*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º"))  +
  plot_layout(ncol  = 1) &
  geom_point()  &
  geom_smooth(method = "lm") & 
  panel_background

```

There is a close spatial relationship between amplitudes and phases of the cEOF1 and TOC planetary wave 1 between `r knitr::combine_words(LatLabel(o3wave_lats))`.
The correlation between the amplitude of both indices is `r wave1_cor[with == "amplitude"]$text` and `r wave1_cor[with == "phase"]$text` between their phases.

## Tropical sources {#tropical}

```{r sst}
sst <- ERSST() %>%
  readRDS() %>% 
  .[season(time) == "SON"] %>%
  normalise_coords() %>%
  .[, w := season_weights(time)] %>% 
  .[, .(t = mean(t*w)), by = .(lon, lat, time = seasonally(time))]
```

```{r cmap}
cmap <- ReadNetCDF(CMAP(), vars = c(pp = "precip"),
                   subset = list(lat = -60:10)) %>%
  normalise_coords() %>%
  .[, w := season_weights(time)] %>% 
  .[, .(pp = mean(pp*w)), by = .(lon, lat, time = seasonally(time))]
```

```{r t2m}
t2m <- ERA5_T2M() %>% 
  ReadNetCDF() %>% 
   normalise_coords() %>%
  .[, w := season_weights(time)] %>% 
  .[, .(t2m = mean(t2m*w)), by = .(lon, lat, time = seasonally(time))]
```

```{r rotated-regressions}
rotate_reim <- function(real, imaginary, angle = 0) {
  z <- complex(real = real, imaginary = imaginary)
  z <- rotate(z, angle)
  list(real = Re(z),
       imaginary = Im(z))
}

angles_regr <- -c(0, 45, 90, 45+90)*pi/180
angles_labs <- setNames(c("Real", "45º", "Imaginary", "135º"), angles_regr*180/pi)

hgt850 <- ERA5_geopotential() %>% 
  ReadNetCDF(vars = c(hgt = "z"), 
             subset = list(level = 850)) %>% 
   normalise_coords() %>%
  .[season(time) == "SON"] %>% 
  .[, w := season_weights(time)] %>% 
  .[, .(hgt = mean(hgt*w)), by = .(lon, lat, time = seasonally(time))]

hgt850 <- ceof[, eof[[1]]$left, by = .(season)] %>%
  .[season == "SON"] %>%
  sep_ReIm(format = "wide") %>%
  .[, hgt := NULL] %>%
  .[, `:=`(Real = as.numeric(scale(Real)), Imaginary = as.numeric(scale(Imaginary))),
    by = .(season, cEOF)] %>%
  .[hgt850, on = "time", allow.cartesian = TRUE]

hgt850_regr <- lapply(angles_regr, function(a) {
  hgt850 %>% 
    copy() %>% 
    .[, c("Real", "Imaginary") := rotate_reim(Real, Imaginary, a)] %>% 
    .[, `:=`(Real = Real/sd(Real), Imaginary = Imaginary/sd(Imaginary)), 
      by = .(season, cEOF)] %>%
    .[, FitLm(hgt, Imaginary, Real, se = TRUE), by = .(lon, lat, cEOF, season)] %>%
    rm_intercept() %>%
    .[term == "Real"] %>% 
    .[, term := factor_ReIm(term)] %>% 
    .[, angle := a*180/pi]
}) %>% 
  rbindlist()

sst_for_regression <- ceof[, eof[[1]]$left, by = .(season)] %>%
  # .[cEOF == "cEOF2"] %>% 
  copy() %>% 
  sep_ReIm(hgt, "wide") %>%
  .[, hgt := NULL] %>%
  .[sst, on = "time", allow.cartesian = TRUE] %>%
  .[season(time) == "SON"] %>% 
  na.omit()

sst_regr <- lapply(angles_regr, function(a) {
  sst_for_regression %>% 
    copy() %>% 
    .[, c("Real", "Imaginary") := rotate_reim(Real, Imaginary, a)] %>% 
    .[, `:=`(Real = Real/sd(Real), Imaginary = Imaginary/sd(Imaginary)), 
      by = .(season, cEOF)] %>%
    .[, FitLm(t, Imaginary, Real, se = TRUE), by = .(lon, lat, cEOF, season)] %>%
    rm_intercept() %>%
    .[term == "Real"] %>% 
    .[, term := factor_ReIm(term)] %>% 
    .[, angle := a*180/pi]
}) %>% 
  rbindlist()

psi_for_regression <-  ceof[, eof[[1]]$left, by = .(season)] %>%
  # .[cEOF == "cEOF2"] %>% 
  copy() %>% 
  sep_ReIm(hgt, "wide") %>%
  .[, hgt := NULL] %>%
  .[psi, on = "time", allow.cartesian = TRUE] %>%
  .[season(time) == "SON"] %>%
  .[lev == 200]

psi_regr <- lapply(angles_regr, function(a) {
  psi_for_regression %>% 
    copy() %>% 
    .[, c("Real", "Imaginary") := rotate_reim(Real, Imaginary, a)] %>% 
    .[, `:=`(Real = Real/sd(Real), Imaginary = Imaginary/sd(Imaginary)), 
      by = .(season, cEOF)] %>%
    .[, FitLm(psi.z, Real, Imaginary, se = TRUE),by = .(lon, lat, lev, cEOF, season)] %>%
    rm_intercept() %>%
    .[term == "Real"] %>% 
    .[, psi.z := Anomaly(estimate), by = .(cEOF, season, lev, lat)] %>%
    .[, c("fx", "fy") := shceof::WaveFlux(.SD, p = 200), 
      by = .(lev, cEOF, season, term)] %>%
    .[, term := factor_ReIm(term)] %>% 
    .[, angle := a*180/pi]
}) %>% 
  rbindlist()

pp_for_regression <- ceof[, eof[[1]]$left, by = season] %>%
  # .[cEOF == "cEOF2"] %>% 
  copy() %>% 
  sep_ReIm(hgt, "wide") %>%
  .[, hgt := NULL] %>%
  .[cmap, on = "time", allow.cartesian = TRUE] %>%
  .[season(time) == "SON"] %>% 
  na.omit() %>% 
  .[, pp := pp/sd(pp), by = .(lon, lat)]

pp_regr <- lapply(angles_regr, function(a) {
  pp_for_regression %>% 
    copy() %>% 
    .[, c("Real", "Imaginary") := rotate_reim(Real, Imaginary, a)] %>% 
    .[, `:=`(Real = Real/sd(Real), Imaginary = Imaginary/sd(Imaginary)), 
      by = .(season, cEOF)] %>%
    .[, FitLm(pp, Imaginary, Real, se = TRUE),
      by = .(lon, lat, cEOF, season)] %>%
    .[term == "Real"] %>% 
    rm_intercept() %>% 
    .[, term := factor_ReIm(term)] %>% 
    .[, angle := a*180/pi]

}) %>%
  rbindlist()

t2m_for_regression <-  ceof[, eof[[1]]$left, by = season] %>%
  # .[cEOF == "cEOF2"] %>% 
  copy() %>% 
  sep_ReIm(hgt, "wide") %>%
  .[, hgt := NULL] %>%
  .[t2m, on = "time", allow.cartesian = TRUE] %>%
  .[season(time) == "SON"] %>% 
  na.omit()

t2m_regr <- lapply(angles_regr, function(a) {
  t2m_for_regression %>% 
    copy() %>% 
    .[, c("Real", "Imaginary") := rotate_reim(Real, Imaginary, a)] %>% 
    .[, `:=`(Real = Real/sd(Real), Imaginary = Imaginary/sd(Imaginary)), 
      by = .(season, cEOF)] %>%
    .[, FitLm(t2m, Imaginary, Real, se = TRUE),
      by = .(lon, lat, cEOF, season)] %>%
    .[term == "Real"] %>% 
    rm_intercept() %>% 
    .[, term := factor_ReIm(term)] %>% 
    .[, angle := a*180/pi]

}) %>%
  rbindlist()
```

```{r enso_cor}
enso_cor <-  ceof[, eof[[1]]$left, by = .(season)] %>%
  sep_ReIm() %>%
  .[, hgt := scale(hgt), by = .(cEOF, part)] %>%
  enso[., on = "time"] %>%
  na.omit() %>%
  .[, correlate(hgt, oni), by = .(cEOF, part)]
```

```{r dmi_cor}
dmi_cor <-  ceof[, eof[[1]]$left, by = .(season)] %>%
  sep_ReIm() %>%
  .[, hgt := scale(hgt), by = .(cEOF, part)] %>%
  dmi[., on = "time"] %>%
  na.omit() %>%
  .[, correlate(hgt, dmi), by = .(cEOF, part)]
```

```{r spoke}
spoke <- function(data, coords) {
  radius <-  grid::unit(.15, "snpc")

  outer_circle <-  grid::circleGrob(x = 0, 
                             y = 0,
                             r = radius*1.2)

  xcentre <- grid::unit(0, "npc") + grid:::grobHeight(outer_circle)/2
  ycentre <-  grid::unit(0, "npc") + grid:::grobWidth(outer_circle)/2

  circle <- grid::circleGrob(x = xcentre, 
                             y = ycentre,
                             r = radius, 
                             gp = grid::gpar(fill = "white", 
                                             alpha = 0.7))
  # browser()
  dx <- cos(coords$angle[1]*pi/180)*grid::grobHeight(circle)/2
  dy <- sin(coords$angle[1]*pi/180)*grid::grobWidth(circle)/2

  line <- grid::segmentsGrob(x0 = xcentre, y0 = ycentre, 
                             x1 = xcentre + dx,
                             y1 = ycentre + dy, 
                             gp = grid::gpar(fill = "black"),
                             arrow = grid::arrow(angle = 13, 
                                                 length =  grid::unit(.15/2, "snpc"),
                                                 type = "closed")
  )

  grid::gTree(children = grid::gList(circle, line))

}
```

```{r plot_sst_psi}
plot_sst_psi <- function(sst_regr, psi_regr) {
  sst_gdata <- sst_regr %>%
    # .[cEOF == "cEOF2"] %>% 
    .[lat %between% c(-90, 10)] %>%
    .[, p_val := Pvaluate(estimate, std.error, df, "fdr"), 
      by = .(cEOF, term, angle)] %>%
    .[, var := "SST"]

  rsst <- max(range(abs(sst_regr[!is.na(cEOF), ]$estimate), na.rm = TRUE))
  breaks_sst <- AnchorBreaks(0, bins = 20, exclude = 0)(c(-rsst, rsst))

  stream_gdata <- psi_regr %>%
    # .[cEOF == "cEOF2"] %>% 
    # .[abs(estimate) > 1, estimate := NA] %>%
    .[lat %between% c(-90, 10)] %>%
    .[, p_val := Pvaluate(estimate, std.error, df, "fdr"), by = .(cEOF, angle, term)] %>%
    .[, var := "Streamfunction"]

  rstream <- max(range(abs(psi_regr$estimate), 
                       na.rm = TRUE))*1e-7
  breaks_stream <- AnchorBreaks(0, bins = 20, exclude = 0)(c(-rstream, rstream))

  # browser()
  stream_gdata %>% 
    periodic(lon = c(0, 360)) %>%
    ggplot(aes(lon, lat)) +

      geom_contour_fill(data = sst_gdata,
                      aes(z = estimate, fill = ..level..),
                      breaks = breaks_sst) +
    geom_contour_pval(data = sst_gdata, aes(z = p_val)) +

    scale_fill_divergent_discretised("SST (K)",
                                     labels = one_less,
                                     guide = guide_colorsteps_bottom(width = 15,
                                                                     show.limits = FALSE, ticks = TRUE,
                                                                     order = 1)) +
    ggnewscale::new_scale_fill() +
    geom_contour_fill(aes(z = estimate*1e-7, fill = ..level..),
                      breaks = breaks_stream) +
    geom_contour_pval(aes(z = p_val)) +
    # geom_streamline(aes(dx = fx, dy = fy), min.L = 3, L = 10,
    #                 size = 0.1, res = 2, nx = 22, ny = 12, 
    #                 arrow.angle = 14, arrow.length = 0.2) +
    geom_vector(aes(dx = fx, dy = fy), skip = 3, min.mag = 0.002, 
                size = 0.1) +
    scale_mag(max_size = .5, guide = "none") +
    geom_qmap(keep = .015) +
    scale_fill_divergent_discretised(latex2exp::TeX("Streamfunction ($m^2s^{-1}\\times 1e^{-7}$)"), 
                                     labels = one_less,
                                     low = scales::muted("red"),
                                     high = scales::muted("blue"),
                                     guide = guide_colorsteps_bottom(15, show.limits = FALSE,
                                                                     order = 99)) +
    grid_panel(spoke, aes(angle = -angle, var = "SST"), inherit.aes = FALSE) +
    scale_x_continuous(NULL, expand = c(0, 0), 
                       breaks = seq(0, 360, by = 60),
                       labels = LonLabel) +

    scale_y_continuous(NULL, breaks  = seq(-90, 0, by = 30), expand = c(0, 0)) +

    facet_grid(angle ~ var, labeller = labeller(angle = angles_labs)) +
    # coord_quickmap(ylim = c(NA, 10)) +
    coord_sf(ylim = c(-89, 10), expand = FALSE,
             # crs = "+proj=longlat +pm=180 +over",
             default_crs = "+proj=longlat +lon_wrap=180 +over") + 
    theme(legend.box = "horizontal",
          legend.text = element_text(size = rel(0.7)), 
          legend.box.spacing = grid::unit(0, "lines") ) +
    tag_facets("rc") 
}
```

```{r  sst-psi-2,  fig.cap = "(ref:sst-psi-2-cap)", fig.width = 7, fig.height=5.6, dependson="plot_sst_psi", fig.env="figure*"}
plot_sst_psi(sst_regr[cEOF == "cEOF2"], psi_regr[cEOF == "cEOF2"]) 
```

(ref:sst-psi-2-cap) Regression maps of cEOF2 with SST (K, column 1) and streamfunction zonal anomalies ($m^2/s\times10^-7$, shaded) with their corresponding activity wave flux (vectors) (column 2) for the 1979 -- 2019 period and different phases of the cEOF (illustrated in the lower-left arrow). Areas marked with dots have p-values smaller than 0.01 adjusted for FDR.

The connections between cEOFs and Tropical sources are also assessed.
Figure \@ref(fig:sst-psi-2) shows the regression maps of Sea Surface Temperatures (SST) and streamfunction anomalies at 200 hPa respectively upon standardised cEOF2, which can be interpreted as the anomalies associated with a 1 standard deviation increase of each index.
Besides showing the regression of the real and imaginary parts, we also show regressions in two intermediate directions.
The Imaginary cEOF2 is associated with strong positive SST anomalies on the Central Pacific and negative anomalies over an area across the North of Australia and New Zealand, the South Pacific Convergence Zone (SPCZ) (Figure \@ref(fig:sst-psi-2).b1).
This pattern is almost canonically positive ENSO [@bamston1997] and indeed, the correlation between the Imaginary cEOF2 and the Oceanic Niño Index [@bamston1997] is significant and very high `r enso_cor[cEOF == "cEOF2" & part == "Imaginary"]$text`.
On top of this ENSO-like pattern, there are positive anomalies in the western Indian Ocean and negative values in the eastern Indian Ocean, resembling a positive Indian Ocean Dipole.
Unsurprisingly, the correlation between the Imaginary cEOF2 and the DMI is `r dmi_cor[cEOF == "cEOF2" & part == "Imaginary"]$text`.

Fittingly, the Imaginary cEOF2 is also associated with strong wave-like streamfunction anomalies emanating from the tropics (Figure \@ref(fig:sst-psi-2).b2), both from the Central Pacific sector and the Incean Ocean.
These are consistent with the effect of ENSO and the Indian Ocean Dipole on the extratropics: the SST anomalies initiate anomalous tropical convection that excites Rossby waves that propagate meridionally towards higher latitudes [@mo2000; @cai2011; @nuncio2015].

Not all phases of the cEOF2 are associated with the same tropical SSTs, however.
Figure \@ref(fig:sst-psi-2).d1 and d2 show that the Real cEOF2 is not associated either with any significant SST nor streamfunction anomalies in the tropics.
The correlation between the Real cEOF2 and ENSO is also not significant (`r enso_cor[cEOF == "cEOF2" & part == "Real"]$text`).
Rows a and c in Fig. 
\@ref(fig:sst-psi-2) show that the intermediate phases are still associated with significant SST anomalies over the Pacific Ocean, but with with slightly different location.
The 135º phase is associated with SST in the central Pacific, while the 45º phase is associated with SST in the eastern Pacific, which correspond roughly to the Eastern Pacific and Central Pacific "flavours" of ENSO (Citas).

(ref:enso-phase-cap) ONI plotted against cEOF2 phase for th 1979 -- 2019 period. Years with magnitude of cEOF2 greater or smaller than the 50th percentile are shown as orange diamonds and green circles and respectively. Black line is the fit ONI \~ sin(phase) computed by weighted OLS using the magnitude of the cEOF2 as weights.

```{r enso-phase, fig.cap = "(ref:enso-phase-cap)", fig.width = width_column, fig.height=3, include = TRUE}
gdata <- ceof[, eof[[1]]$left, by = .(season)] %>%
  .[cEOF == "cEOF2"] %>% 
  # .[, hgt := Anomaly(hgt)] %>% 
  .[, arg := Arg(hgt)] %>%
  .[, mag := Mod(hgt)] %>%
  .[enso, on = "time"] %>%
  na.omit() %>%
  .[, q := ifelse(mag >= quantile(mag, 0.5), "|cEOF2| > 50%", "|cEOF2| < 50%")]

cor_enso_mag <- gdata[, correlate(mag, abs(oni))]

cor_enso_spearman <- gdata[, cor.test(mag, abs(oni), method = "spearman")]

cor_enso_mag_outlier <- gdata[oni < 1.5][, correlate(mag, abs(oni))]

cor_enso_mag_quantile <- gdata[mag >= quantile(mag, 0.5)] %>% 
  .[oni < 1.5] %>% 
  .[, correlate(mag, abs(oni))]

cos_model <- gdata %>% 
  .[, lm(oni ~ I(sin(arg)) - 1, weights = mag )]

equation <- latex2exp::TeX(paste0("$\\mathrm{ONI} = ", 
                                  signif(coef(cos_model)[1], 2), "\\sin{(\\mathrm{phase})}$"))

gdata %>% 
  ggplot(aes(arg, oni)) +
  geom_hline(yintercept = c(-0.5, 0, 0.5), size = 0.2, color = "gray50") +
  geom_vline(xintercept = 0, size = 0.2, color = "gray50") +
  annotate(shadowtext:::GeomShadowText,
           label = c("- Real", "- Imaginary", "+ Real", "+ Imaginary", "- Real"),
           color = "gray60", bg.colour = "white", bg.r = 0.2,
           size = 3,
           y = -Inf, x = c(-pi, -pi/2, 0, pi/2, pi), angle = 90, hjust = 0) +

  geom_smooth(data = ~qwrap(.x, arg = c(-pi, pi) ~ c(-2*pi, 2*pi)),
              method = "lm", aes(weight = mag),
              formula = y ~ I(sin(x)) -1,
              se = FALSE, color = "black", size = 0.5) +
  geom_point(aes(color = q,
                 shape  = q),
             size = 3) +
  ggrepel::geom_text_repel(data = ~.x[order(-oni)][1:3],
                           aes(label = year(time))) +

  scale_color_brewer("", palette = "Dark2") +
  scale_shape_manual("", values = c(20, 18)) +
  scale_y_continuous("ONI") +
  scale_x_continuous("Phase",
                     breaks = seq(-pi, pi, by = 45*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  coord_cartesian(xlim = c(-pi, pi)) +
  labs(subtitle = equation) +
  panel_background
```

To better explore the relationship between tropical forcing and phase of the cEOF2, Figure \@ref(fig:enso-phase) shows the ONI index against the cEOF2 phase for each SON trimester between 1979 and 2019, highlighting years in which the magnitude of cEOF2 is above the median.
In years with positive ONI, the phase of the cEOF2 is mostly around +90º (corresponding with positive imaginary part) and vice versa.
In years with near neutral ENSO, the phase of the cEOF2 is much more variable.
The black line in Figure \@ref(fig:enso-phase) is a sinusoidal fit of the relationship between ONI and cEOF2 phase.
The $r^2$ corresponding to the fit is `r signif(summary(cos_model)$r.squared, 2)`, statistically significant with p-value `r report_p(coef(summary(cos_model))[4])`, indicating a quasi-sinusoidal relation between these two variables.

The correlation between the absolute magnitude of the ONI and the magnitude of the cEOF2 is `r cor_enso_mag$text`.
This relationship, however, appears to be driven only by the three years with strongest ENSO events in the period (`r knitr::combine_words(year(gdata[order(-oni)][1:3]$time))`) which also coincide with the three years with strongest cEOF2 magnitude.
If those years are removed, the correlation becomes non-significant (`r cor_enso_mag_outlier$text`).
Furthermore, even when using all years, the Spearman correlation --which is robust to outliers-- is also non-significant (`r signif(cor_enso_spearman$estimate, 2)`, p-value`r report_p(cor_enso_spearman$p.value)`).
Therefore, although the location of tropical SST seem to have an effect in defining the phase of the cEOF2, the relationship between the magnitude of the cEOF2 train wave and ONI remains uncertain.

It could be concluded that the wave train represented by cEOF2 can be both part of the internal variability of the extratropical atmosphere or forced by tropical SSTs.
In the former case, the wave train has little phase preference.
However, when cEOF2 is excited by tropical SST variability, it tends to remain locked to the imaginary phase.
This explains the relative over-abundance of years with cEOF2 near positive and negative imaginary phase in Figure \@ref(fig:phase-histogram).

Unlike the cEOF2 case, there is no significant pattern of SST anomalies associated with either the Real or Imaginary cEOF1 (Figure \@ref(fig:sst-psi-1)).
Consistently, streamfunction anomalies do not show any tropical influence.
Instead, the real and Imaginary cEOF1 are associated with zonally wave activity fluxes in the extra-tropics around 60ºS, except for an equatorward flow from the coast of Antarctica around 150ºE in the real component.
This suggests that its variability is driven primary by the internal variability of the extra-tropics.

## Precipitation and 2-metre temperature {#precipitation}

```{r pp-t2m-r2, fig.cap = "(ref:cap-pp-t2m-r2)", fig.width=3.3*2, fig.height=3.3}
breaks <- seq(.2, 1, by = .1)

rbind(Precipitation = pp_regr,
      `2-metre Temperature` = t2m_regr, idcol = "var") %>% 
  .[angle == 0] %>% 
  .[is.finite(r.squared)] %>% 
  # .[cEOF == "cEOF2"] %>% 
  periodic(lon = c(0, 360)) %>%
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = r.squared, fill = ..level..), breaks = breaks) +
  geom_contour_tanaka(aes(z = r.squared), breaks = breaks) +
  # geom_text_contour(aes(z = r.squared, 
  #                       label = ifelse(..level.. > 0.2, scales::percent(..level..), NA)), 
  #                   breaks = breaks, size = 2,
  #                   skip = 1, stroke = 0.15, stroke.colour = "white") +
  geom_qmap(keep = .015) +
  scale_fill_divergent_discretised(name = "Explained variance", midpoint = 0,
                       labels = scales::percent_format(),
                       guide = guide_colorsteps_bottom(15, show.limits = TRUE)) +

  scale_x_continuous(NULL, expand = c(0, 0),
                     breaks = seq(0, 360, by = 60),
                     labels = LonLabel) +
  scale_y_continuous(NULL, breaks  = seq(-90, 0, by = 30), expand = c(0, 0)) +

  facet_grid(var ~ cEOF) +
  # coord_quickmap(ylim = c(NA, 10)) +
  coord_sf(ylim = c(NA, 10),
           default_crs = "+proj=longlat +lon_wrap=180 +over") + 
  theme(legend.box = "horizontal",
        legend.text = element_text(size = rel(0.7)), 
        legend.box.spacing = grid::unit(0, "lines") ) +
  tag_facets("rc") 
```

(ref:cap-pp-t2m-r2) Explained variance ($r^2$) of 2-metre temperature (row a) and precipitation (row b) with cEOF1 (column 1) and cEOF2 (column 2).

The influence of cEOFs variability in continental rainfall and 2-metre temperature in the Southern Hemisphere is also explored.
Figure \@ref(fig:pp-t2m-r2) shows the $r^2$ between the cEOF1 and cEOF2 and 2-metre temperature and precipitation.
For the cEOF1 there are no significant signals in precipitation and most of the temperature signal is low and restricted to the eastern coast of Antarctica.

There is a strong signal between both variables and cEOF2, with explained variance greater than 50% in some regions.
For 2-metre temperature, there are high values in the tropical Pacific and the SPCZ, as well as the region following an arc between New Zealand and the South Atlantic.
Over the continents, moderate values of about 30% variance explained can be seen in southern Australia, some regions of South America and the Antarctic Peninsula.
For precipitation, there are high values over the tropics.
At higher latitudes, moderate values are observed over eastern Australia and some regions of southern South America.

Since the cEOF1 has a relatively weak signal, we will focus on the cEOF2.

```{r add_continents}
continents <- list(africa  = list(lon = c(0, 60),
                                  lat = c(-40, 0)),
                   oceania = list(lon = c(100, 180),
                                  lat = c(-50, -13)),
                   america = list(lon = c(265, 325),
                                  lat = c(-60, -10)))

add_continent <- function(dt) {
  dt[, continent := fcase(lon %between% continents$africa$lon &
                            lat %between% continents$africa$lat, "africa",
                          lon %between% continents$oceania$lon &
                            lat %between% continents$oceania$lat, "oceania",
                          lon %between% continents$america$lon &
                            lat %between% continents$america$lat, "america")] %>%
    .[, continent := factor(continent, levels = c("africa", "oceania", "america"), ordered = TRUE)] %>%
    .[]
}

crop_continent <- function(continent) {
  bbox <- c(xmin = min(continents[[continent]]$lon), 
            xmax = max(continents[[continent]]$lon),
            ymin = min(continents[[continent]]$lat),
            ymax = max(continents[[continent]]$lat))

  function(data) {
    s2 <- sf::sf_use_s2(FALSE)
    on.exit(sf::sf_use_s2(s2))
    data <- sf::st_crop(data, bbox)
    data$continent <- continent
    data
  }
}
```

```{r plot_pp}
plot_pp <- function(pp_regr, t2m_regr, hgt850_regr) {
  lats <- c(-55, 10)
  breaks_pp <- AnchorBreaks(exclude = 0)(c(-.9, .9))

  pp_regr <- pp_regr %>% 
    copy() %>% 
    .[, var := "Precipitation"] %>% 
        .[, p.value := Pvaluate(estimate, std.error, df, "none"), 
      by = .(cEOF, angle)]

  t2m_regr <- t2m_regr %>% 
    copy() %>% 
    .[, var := "2m Temperature"] %>% 
    .[, p.value := Pvaluate(estimate, std.error, df, "none"), by = .(cEOF, angle)]

  pp_regr %>%
    # .[lat %between% lats] %>%

    periodic(lon = c(0, 360)) %>%
    ggplot(aes(lon, lat)) +

    geom_contour_fill(aes(z = estimate, fill = ..level..),
                      breaks = breaks_pp) +
    geom_contour_tanaka(aes(z = estimate), breaks = breaks_pp) +

    scale_fill_divergent_discretised("Precipitation (correlation)", 
                                     # limits = c(-1, 1),
                                     low = "#8E521C",
                                     high = "#00665E",
                                     guide = guide_colorsteps_bottom(15, order = 99)) +
    geom_contour_pval(aes(z = p.value)) +

    ggnewscale::new_scale_fill() +

    geom_contour_fill(data = t2m_regr, 
                      aes(z = estimate, fill = ..level..), 
                      breaks = AnchorBreaks(exclude = 0)) +
    geom_contour_tanaka(data = t2m_regr, 
                        aes(z = estimate),
                        breaks = AnchorBreaks(exclude = 0)) +

    scale_fill_divergent_discretised("Temperature",
                                     # limits = c(-1, 1),
                                     guide = guide_colorsteps_bottom(15, order = 1)) +
    geom_contour_pval(data = t2m_regr, aes(z = p.value)) +

    geom_contour2(data = copy(hgt850_regr)[, var := "2m Temperature"], 
                  aes(z = estimate, linetype = factor(-sign(..level..))), 
                  size = 0.2, breaks = AnchorBreaks(0, exclude = 0)) +

    geom_qmap() +

    grid_panel(spoke, aes(angle = -angle, var = "SST"), inherit.aes = FALSE) +

    scale_x_continuous("", expand = c(0, 0), breaks = seq(0, 360, by = 30), 
                       labels = LonLabel) +
    scale_y_continuous("", expand = c(0, 0), breaks = seq(-90, 0, by = 15), 
                       labels = LatLabel) +
    # scale_y_latitude(ticks = 15, labels = LatLabel) +
    scale_linetype_discrete(guide = "none") +
    facet_grid(angle ~ var, labeller = labeller(angle = angles_labs)) +
    axis_labs_smol +
    tag_facets("rc") +
    coord_sf(ylim = c(NA, 10),
             default_crs = "+proj=longlat +over") +
    theme(panel.spacing = grid::unit(1, "lines")) +
    theme(legend.box = "horizontal",
          legend.text = element_text(size = rel(0.7)), 
          legend.box.spacing = grid::unit(0, "lines") )

}
```

```{r boxes}
sesa <- list(lon = c(300, 315),
             lat = c(-35, -22))
box_australia <- data.frame(cEOF = "cEOF2", 
                            lon = c(130, 155),
                            lat = c(-40, -15))
```

Figure \@ref(fig:pp-temp-2) shows regression maps of 2-meter temperature (column 1) and precipitation (column 2) upon standardised cEOF2 which, again, can be interpreted as the anomalies associates with a 1 standard deviation increase.

(ref:pp-temp-cap) Regression of SON mean 2-meter temperature (K, shaded) and 850hPa geopotential height (m, contours) (column 1), and precipitation (correlation, column 2) with different phases of cEOF2. For the 1979 -- 2019. Areas marked with dots have p-values smaller than 0.01 adjusted for False Detection Rate.

```{r pp-temp-2, fig.cap="(ref:pp-temp-cap)", fig.width=6.6, fig.height=5.5, dependson="plot_pp"}
plot_pp(pp_regr[cEOF == "cEOF2"], t2m_regr[cEOF == "cEOF2"], hgt850_regr[cEOF == "cEOF2"])
```

Temperature anomalies associated with the Imaginary cEOF (Fig. \@ref(fig:pp-temp-2).b1) show positive values in the tropical Pacific, consistent with SSTs anomalies associated with the same phase (Fig. \@ref(fig:sst-psi-2).b1).

At higher latitudes there is a wave-like pattern of positive and negative values that coincide with the nodes of the 850hPa geopotential height regression patterns.
This is consistent with temperature anomalies produced by meridional advection of temperature by the meridional winds arising from geostrophic balance.

Over the continents, the Imaginary cEOF2 (Fig. \@ref(fig:pp-temp-2)b.1) is associated with positive temperatures in southern Australia and negative temperature in southern South America and the Antarctic Peninsula.

The temperatures anomalies associated with the Real cEOF2 (Fig. \@ref(fig:pp-temp-2)d.1) are less extensive.
There is no significant signal over the continents except for positive anomalies near the Antarctic Peninsula.

Tropical precipitation anomalies associated with the Imaginary cEOF2 are strong, with positive anomalies in the central Pacific and western Indian, an negative anomalies in the eastern Pacific.
This pattern is consistent with the SST anomalies shown before as the increased SST enhances tropical convection and vice versa.

On the extra-tropics, the Imaginary cEOFs is correlated with drier conditions over eastern Australia; a similar signal as the one associated with ENSO @cai2011.
However, the Imaginary cEOF2 is not the direction most correlated with precipitation in that area.
The 135º direction (positive Imaginary + negative Real cEOF2) is associated with stronger and more extensive correlations with precipitation over Australia and New Zealand.
This Australian signal is most likely there due to the more direct impacts of ENSO rather than on the interconnection pattern represented by the cEOF2.

(ref:australia-pp-phase-cap) Same as Figure \@ref(fig:enso-phase) but for precipitation anomalies averaged between `r knitr::combine_words(LonLabel(box_australia$lon))` and `r knitr::combine_words(LatLabel(box_australia$lat))` (box shown in Figure \@ref(fig:pp-oceania) column b) in $\frac{\mathrm{mm}}{\mathrm{day}}$. Black line is the fit precipitation \~ sin(phase) + cos(phase) computed by weighted OLS using the magnitude of the cEOF2 as weights.

```{r australia-pp-phase, fig.cap = "(ref:australia-pp-phase-cap)", fig.width = width_column, fig.height=3, include = FALSE}
gdata <- cmap %>% 
  .[lon %between% box_australia$lon  & lat %between% box_australia$lat] %>% 
  .[, .(pp = weighted.mean(pp, cos(lat*pi/180))), by = time] %>% 
  .[, pp := Anomaly(pp), by = season(time)] %>% 
  .[ceof$eof[[1]]$left, on  = "time"] %>% 
  .[cEOF == "cEOF2"] %>% 
  .[, arg := Arg(hgt)] %>% 
  .[, mag := Mod(hgt)] %>% 
  .[, q := ifelse(mag >= quantile(mag, 0.5), "|cEOF2| > 50%", "|cEOF2| < 50%")]

model <- lm(pp ~ I(sin(arg)) + I(cos(arg)) - 1, weights = mag, data = gdata)
fstatistic <- summary(model)$fstatistic
p_value <- pf(fstatistic["value"], fstatistic["numdf"], fstatistic["dendf"], 
              lower.tail = FALSE)

equation <- latex2exp::TeX(paste0("$\\mathrm{Precipitation}_{\\mathrm{East Australia}} = ", 
                                  signif(coef(model)[1], 2), "\\sin{(\\mathrm{phase})} + ", 
                                  signif(coef(model)[2], 2), "\\cos{(\\mathrm{phase})}$"))

gdata %>% 
  ggplot(aes(Arg(hgt), pp)) +
  geom_hline(yintercept = 0, size = 0.2, color = "gray50") +
  geom_vline(xintercept = 0, size = 0.2, color = "gray50") +
  annotate(shadowtext:::GeomShadowText,
           label = c("- Real", "- Imaginary", "+ Real", "+ Imaginary", "- Real"),
           color = "gray60", bg.colour = "white", bg.r = 0.2,
           size = 3,
           y = -Inf, x = c(-pi, -pi/2, 0, pi/2, pi), angle = 90, hjust = 0) +

  geom_smooth(data = ~qwrap(.x, arg = c(-pi, pi) ~ c(-2*pi, 2*pi)),
              method = "lm", aes(weight = mag),
              formula = y ~ I(sin(x)) + I(cos(x)) -1,
              se = FALSE, color = "black", size = 0.5, 
              fullrange = TRUE) +
  geom_point(aes(colour = q,
                 shape  = q),
             size = 3) +

  # annotate("text", x = 90*pi/180, y = 1, label = equation, parse = TRUE) +

  scale_color_brewer(NULL, palette = "Dark2") +
  scale_shape_manual(NULL, values = c(20, 18)) +
  scale_y_continuous("Precipitation anomaly") +
  scale_x_continuous("Phase",
                     breaks = seq(-pi, pi, by = 30*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  coord_cartesian(xlim = c(-pi, pi)) +
  labs(subtitle = equation) +
  theme(plot.subtitle = element_text(size = 8)) +
  panel_background
```

<!-- To better understand the relationship between precipitation anomalies over eastern Australia and the cEOF2, Figure @ref(fig:australia-pp-phase) plots precipitation anomalies averaged over eastern Australia (box in Figure @ref(fig:pp-oceania)) as a function of cEOF2 phase. -->

<!-- The black line shows a sinusoidal fit, whose coefficients are fitted by weighted least squares using the magnitude of the EFO2 as weight. -->

<!-- The $r^2$ corresponding to the fit is `r signif(summary(model)$r.squared, 2)` (p-value `r report_p(p_value)`). -->

<!-- There is a strong relationship between the phase of the cEOF2 and precipitation anomalies in eastern Australia, but the Real and Imaginary phases we chose are not exactly aligned with the direction that maximises this relationship. -->

Over South America, the Imaginary cEOF has positive correlations with precipitation in South Eastern South America (SESA) and central Chile, and negative correlations in eastern Brazil.
This again is a well-known springtime precipitation signature of ENSO [@cai2020a] and it is also similar to the precipitation anomalies associated with the A-SAM [\@ @campitelli2022].

(ref:pp-america-cap) As Figure \@ref(fig:pp-oceania) but for South America

```{r pp-america, fig.cap = "(ref:pp-america-cap)", fig.width=width_column, fig.height=width_column*1.1}
# plot_pp(data = regr_pp, which_cEOF = c("cEOF1", "cEOF2"), which_continent = "america")  +
#   scale_fill_divergent_discretised("Precipitation (mm/day)", 
#                                    limits = c(-1, 1),
#                                    # oob = scales::squish,
#                                    low = "#8E521C",
#                                    high = "#00665E",
#                                    guide = guide_colorsteps_bottom(15)) 
# annotate("rect", xmin = min(sesa$lon), xmax = max(sesa$lon), 
#          ymin = min(sesa$lat), ymax = max(sesa$lat),
#          fill = NA, color = "gray50", size = 0.3)
#          
```

```{r sinusoidal-sesa}
gdata <- cmap %>% 
  .[lon %between% sesa$lon  & lat %between% sesa$lat] %>% 
  .[, .(pp = weighted.mean(pp, cos(lat*pi/180))), by = time] %>% 
  .[, pp := Anomaly(pp), by = season(time)] %>% 
  .[ceof$eof[[1]]$left, on  = "time"] %>% 
  .[cEOF == "cEOF2"] %>% 
  .[, arg := Arg(hgt)] %>% 
  .[, mag := Mod(hgt)]

model <- lm(pp ~ I(sin(arg)) + I(cos(arg)) - 1, data = gdata)

fstatistic <- summary(model)$fstatistic
p_value <- pf(fstatistic["value"], fstatistic["numdf"], fstatistic["dendf"], 
              lower.tail = FALSE)
```

This is not surprising considering the close relationship between the ONI, the A-SAM index and the Imaginary cEOF2 shown previously, but further consolidates the identification of this mode with the PSA pattern.
Resembling the relationship between ONI and the phase of cEOF2 (Fig. \@ref(fig:enso-phase)), there is a cEOF2 phase dependence of the precipitation anomalies in SESA (not shown).
These variables could be related with a significant sinusoidal fit that has a statistically significant coefficient of determination of `r round(summary(model)$r.squared, 2)`.

The precipitation signal associated with the Real cEOF2 is less strong.
There is a residual positive correlation in the equatorial eastern Pacific and small, not statistically significant positive correlations over eastern Australia and negative ones over New Zealand.

# Conclusions

In this study we assessed the Austral Spring Extratropical Southern Hemisphere zonally asymmetric circulation.
For this purpose, two complex indices were derived using Complex Empirical Orthogonal Functions.
This allowed us to characterise both the amplitude and phase of planetary waves that are not perfectly organised as sinusoidal waves with constant amplitude at each latitude circle.

The first complex EOF represents the variability of the zonal wave 1 in the stratosphere, but represents a more zonally symmetric monopole in the troposphere.
There is a statistically positive trend in the magnitude of this cEOF, which is consistent with previous studies that showed secular changes in springtime wave-1-like patterns [e.g. @raphael2003].
This mode is closely related to stratospheric variability such as anomalies in Total Ozone Column.
Otherwise, this complex EOF is not related with SST variability and continental precipitation in the Southern Hemisphere.

The second complex EOF represents a wave-3 pattern with maximum magnitude in the Pacific sector.
Essentially, it is an alternate representation of the PSA1 and PSA2 patterns [@mo2001].
We show that the Imaginary cEOF2 can be roughly identified with the PSA1 and the Real cEOF2 with the PSA2.

There is some disagreement in the literature of whether the phase of the PSA pattern is affected by the location of tropical SST anomalies [@irving2016].
We show not only that the cEOF2 tends to be in the imaginary phase (\~PSA1) when the ENSO region is warm, but that central Pacific SST anomalies tend to move the cEOF towards the negative real phase and eastern Pacific SST anomalies tend to move it towards positive real phase.
When ENSO is neutral, the cEOF2 is still active, but with no particular preferred phase.
This mirrors the results of @cai2002, who showed that the CSIRO Model can create PSA-like variability even in the absence of ENSO forcing (with a climatological run), but the variability of one of the PSA modes was enhanced when adding the ENSO signal.
The magnitude of the cEOF2 appears not to be related to the magnitude of ENSO, except for the three strongest ENSO events in the period which coincide with the three strongest cEOF2 years.

The sensitivity of the phase of the cEOF2 to the location of the tropical SST anomalies was also seen by @ciasto2015, who detected similar Rossby wave patterns associated with central Pacific and eastern Pacific SST anomalies but with a change in phase.

We further show that the Imaginary cEOF2 is closely related to the Southern Annular Mode in the troposphere.
In fact, it has a very close resemblance to the zonally asymmetric portion of the SAM.
This raises the possibility that the asymmetric component of the SAM is actually a statistical contamination of the PSA mode.

Precipitation anomalies in South America associated with the Imaginary cEOF2 show a clear ENSO-like impact, with positive anomalies in South-Eastern South America, negative anomalies in Southern Brazil and positive anomalies in central Chile.
Precipitation anomalies associated with the Real cEOF2 are low and not statistically significant, showing that the Imaginary phase is optimally aligned with the direction of maximum precipitation impacts.

On the other hand, over Australia, both the Real and Imaginary phase are associated with significant precipitation anomalies, and we further show that the direction of maximum impact is not aligned with our chosen rotation of cEOF2.
However, this underscores the benefit of using complex EOF, since it would be trivial to rotate it.
That is, look for the rotation of complex EOF that maximises the relationship and use that as the relevant index for study or forecast.

Some aspects of our work are in direct contrast with @irving2016 analysis of the PSA.
They didn't find neither a strong association between PSA-like variability and ENSO, nor a relationship between it and Australian precipitation.
Part of the disagreement could be due to the fact that we focus data only for SON, which is when the teleconnection pattern is strongest, instead of the whole year.
@irving2016 use of ERA5 forectasted precipitation instead of the more observationally constrained CMAP dataset might also be an issue.
But in light of the previous discussion on the importance of the phase of the cEOF2 in its relationship with ENSO and with precipitation impacts, a big difference could be fact that @irving2016's methodology is insensitive to phase changes of the PSA-like variability.

To the extent that the cEOF2 can be identified with the PSA, what is the advantage of the technique proposed here?
The spatial fields that the cEOF2 index describes are in quadrature by construction, which makes it possible to derive a proper amplitude and phase.
Furthermore, because of the way they are constructed, they are not forced to be orthogonal to the SAM.
This makes it possible to study their relationship with the SAM free of the statistical contamination that comes with the usual definition of the PSA as the second and third EOFs.

Further work should extend this analysis to other seasons and further study the relationship between the cEOF2 and the SAM.
In particular, it is not clear whether the close identification between the Imaginary cEOF2 and the Asymmetric SAM arises from physical processes or statistical contamination.

## Code availability {.unnumbered}

A version-controlled repository of the code used to create this analysis, including the code used to download the data can be found at <https://github.com/eliocamp/shceof>.

-   The Oceanic Niño Index is available via NOAA's Climate Prediction Center: <https://www.cpc.ncep.noaa.gov/products/analysis_monitoring/ensostuff/detrend.nino34.ascii.txt>.

-   The Dipole Mode Index is available via Global Climate Observing System Working Group on Surface Pressure: <https://psl.noaa.gov/gcos_wgsp/Timeseries/Data/dmi.had.long.data>

\appendix

\counterwithin{figure}{section}

# Extra figures

\newpage

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

(ref:sst-psi-1-cap) Same as Figure \@ref(fig:sst-psi-2) but for cEOF1.

```{r sst-psi-1, fig.cap = "(ref:sst-psi-1-cap)", fig.width = 7, fig.height=5.6, dependson="plot_sst_psi", fig.env = "figure*"}
plot_sst_psi(sst_regr[cEOF == "cEOF1"], psi_regr[cEOF == "cEOF1"]) 
```

(ref:pp-temp-cap-1) Same as Figure \@ref(fig:pp-temp-2) but for cEOF1.

(ref:pp-temp-cap-1) Same as Figure \@ref(fig:pp-temp-2) but for cEOF1.

```{r pp-temp-1, fig.cap="(ref:pp-temp-cap-1)", fig.width=6.6, fig.height=5.5, dependson="plot_pp"}
plot_pp(pp_regr[cEOF == "cEOF1"], t2m_regr[cEOF == "cEOF1"], hgt850_regr[cEOF == "cEOF1"])
```